\documentclass[12pt]{article}

% House style preamble
\input{.house-style/preamble.tex}
\usepackage{tikz}
\usetikzlibrary{arrows.meta,positioning,calc}
\usepackage{placeins}
\fancyhead[L]{\small\scshape Delegation Assurance}

\setlength{\headheight}{25.3pt}
\emergencystretch=2em

\hypersetup{
  pdftitle={Delegation Assurance for AI Systems: Typed Authorization Semantics and Comparative Evaluation},
  pdfkeywords={AI security, AI agents, delegation assurance, prompt injection, auditability, safety cases, LLM judges}
}

\title{Delegation Assurance for AI Systems: Typed Authorization Semantics and Comparative Evaluation}
\author{Brett Reynolds \orcidlink{0000-0003-0073-7195}%
\thanks{I used OpenAI Codex and Anthropic Claude extensively in drafting and revising the paper. I reviewed, edited, and approved all the material and take full responsibility for the final text and conclusions. \href{mailto:brett.reynolds@humber.ca}{brett.reynolds@humber.ca}}\\
Humber Polytechnic \& University of Toronto}
\date{July 2026}

\begin{document}
\maketitle

\begin{abstract}
Current agent-security work provides threat taxonomies, attacks, defences, benchmarks, and architectural controls. Less developed is the authorization question: how should an evaluator represent and test whether an operation or bounded trajectory executed through an AI deployment remained within delegated authority? I use \term{delegation assurance} for regime-relative authorization semantics, typed trace representations, failure diagnosis, and three distinct comparative programmes. The account treats a single action as the \(n=1\) trajectory case and authorization standing as temporal and operation-specific. It separates the normative proposition from the record that represents it, the findings an evaluator produces, the proposed use of those findings, any extension to unobserved cases, and any causal explanation. It also distinguishes a closed-world stipulated regime from an institutional regime whose identity and applicability may require adjudication. For language-governed action, the framework separates independently established contextual status and standing from recorded assignments, conflict rules, licensing decisions, proposals, gate decisions, and execution. A versioned closed-world engine derives grant validity, transition effects, action authorization, and trajectory conditions from reference-free traces; separate hidden oracles test those derivations without entering reviewer or predictor views. Frozen machine-readable specifications and an analyzer implement local-discrimination, predictive-ablation, and reviewer-reconstruction decision rules, although no target-study outcomes are reported. Records can support reconstruction without proving institutional authority; randomized representations can identify a local package effect without randomizing normative standing; and neither result establishes performance on a new regime, reviewer population, system, or time. The companion evidentiary-assurance account asks what a preserved record warrants under a declared review procedure.
\end{abstract}

\noindent\textbf{Keywords:} AI security; AI agents; delegation assurance; prompt injection; auditability; safety cases; LLM judges

\input{sections-delegation/01-introduction}
\input{sections-delegation/02-existing-work}
\input{sections-delegation/03-framework}
\input{sections-delegation/04-failure-evidence-mapping}
\input{sections-delegation/05-language-mediated-authority}
\input{sections-delegation/06-implications}
\input{sections-delegation/07-compositional-delegation}
\clearpage
\input{sections-delegation/07-conclusion}

\clearpage
\begingroup
\renewcommand*{\bibfont}{\small}
\setlength{\bibitemsep}{0.25\baselineskip}
\printbibliography
\endgroup

\end{document}
