\documentclass[12pt]{article}

% House style preamble
\input{.house-style/preamble.tex}
\usepackage{tikz}
\usetikzlibrary{arrows.meta,positioning}
\usepackage{longtable}
\usepackage[section]{placeins}
\usepackage{flafter}
\fancyhead[L]{\small\scshape Evidentiary Assurance}

\setlength{\headheight}{25.3pt}
\emergencystretch=2em

\hypersetup{
  pdftitle={Evidentiary Assurance for Delegated Authority in AI Systems},
  pdfkeywords={AI governance, AI evidence, delegated authority, answerability, record adequacy, assurance cases, assessment validity, electronic evidence}
}

\title{Evidentiary Assurance for\\Delegated Authority in AI Systems}
\author{Brett Reynolds \orcidlink{0000-0003-0073-7195}%
\thanks{I used OpenAI Codex and Anthropic Claude extensively in drafting and revising the paper. I reviewed, edited, and approved all the material and take full responsibility for the final text and conclusions. \href{mailto:brett.reynolds@humber.ca}{brett.reynolds@humber.ca}}\\
Humber Polytechnic \& University of Toronto}
\date{July 2026}

\begin{document}
\maketitle

\begin{abstract}
AI governance increasingly requires logs, documentation, human oversight, impact assessment, provenance, recourse, and auditability. These requirements generate evidence without answering the question a contested case actually turns on: what does this record warrant about whether this action was authorized? A bad record doesn't make an authorized action unauthorized, and a pristine record can document an invalid delegation, so the historical question and the evidentiary one have to be answered separately. This paper supplies the evidentiary half. It keeps four review outputs distinct rather than merging them into an assurance score: what the record warrants about authorization, whether the record is adequate for the declared review, whether any single person or organization is identifiable, dutied, and reachable enough to answer, and whether any single forum has the access, competence, independence, and remedial power to review the action. The last two are evaluated of named candidates, because properties satisfied piecemeal across different parties describe no bearer and no route. Verdicts separate an unevaluable record from an adequate record whose showing falls short, and both from affirmative evidence of failure, since those call for different remedies. Applicability is fixed before the record is opened. An executable artifact encodes the distinctions and tries to break them with matched and adversarial fixtures; because no reviewer has yet answered them, the procedure remains a research hypothesis rather than a validated instrument.
\end{abstract}

\noindent\textbf{Keywords:} AI governance; AI evidence; delegated authority; answerability; record adequacy; assurance cases; assessment validity; electronic evidence

\input{sections-evidentiary/01-introduction}
\input{sections-evidentiary/02-evidence-primitives}
\input{sections-evidentiary/03-governance-law}
\input{sections-evidentiary/04-technical-mechanisms}
\input{sections-evidentiary/05-sufficiency-test}
\input{sections-evidentiary/05-research-program}
\clearpage
\input{sections-evidentiary/06-conclusion}
\clearpage
\input{sections-evidentiary/07-appendix-fixture}

\clearpage
\begingroup
\renewcommand*{\bibfont}{\footnotesize}
\setlength{\bibitemsep}{0.25\baselineskip}
\printbibliography
\endgroup

\end{document}
