\section{The Evaluation Problem}
\label{sec:evaluation-problem}

Frontier-model safety evaluation increasingly depends on language-mediated control. A useful evaluation asks whether the model can solve the task, identify which natural-language material has directive force, apply the relevant policy boundary, and interpret what a transcript shows about the model's behaviour.

This connects AI safety evaluation to \term{language-mediated delegation}: systems act through natural-language instructions, tool outputs, retrieved documents, policy text, and agent transcripts. A safety evaluation is incomplete if it only asks whether the final output was acceptable. It also has to ask whether the system preserved control over which language counted as an instruction, whose authority it carried, what action it licensed, and how any failure should be attributed.

The problem already appears in work on holistic language-model evaluation, which treats model assessment as a multi-scenario, multi-metric problem rather than a single accuracy number \parencite{liang2022holistic}. Safety evaluation adds a sharper version of the same problem: the relevant behaviour is often not just an answer, but how the model responds to an instruction, policy, or source relation.

The practical bottleneck is visible in prompt injection, model-policy evaluation, red-team triage, LLM-as-judge grading, and agent transcript review. Prompt-injection and indirect-prompt-injection research show that LLM-integrated applications can blur the boundary between data and instructions across realistic application settings \parencite{greshake2023not,yi2025bipia}. SEP formalizes one central version of that problem: whether a model executes an embedded probe or processes it as data \parencite{zverev2025separate}. Instruction-hierarchy work frames a neighbouring vulnerability as failure to follow privileged rather than lower-priority instructions \parencite{wallace2024instruction}. Agent evaluations make the issue harder because untrusted content can arrive through tools and retrieved documents rather than through the user turn itself \parencite{debenedetti2024agentdojo}.

Three questions have to remain separate. \term{authorization} asks which source is entitled to direct an action (authorization standing, in the delegation-assurance sense \parencite{reynolds2026delegationAssurance}, rather than the licensed-action verdict that account calls action authorization); \term{causal sensitivity} asks which contextual change actually alters the response; and \term{projection} asks where the measured relation is expected to recur. Context can change behaviour without being authorized to guide it, and an authorized source can fail to affect behaviour. When a declared projection concerns the same causal query in a different domain, it raises the further, query-specific problem of transportability \parencite{pearlBareinboim2014externalValidity}.

In each setting, a surface string may be an instruction, a quotation, a cited passage, a tool output, a policy example, a user request, or an adversarial attempt to change authority. A pass/fail label loses information when it doesn't say which role the string had and why.

Evaluation labels function here as inference licenses, but a single label can conceal four different bearers. \term{Reference projectibility} concerns the licensed-action relation generated by a frozen stipulated-regime template and output-blind adjudication procedure. \term{Behavioural projectibility} concerns a versioned evaluation-configuration family, including its enumerated wrappers and context-assembly rules. \term{Measurement projectibility} concerns a versioned scorer, codebook, implementation, adjudication procedure, criterion semantics, and predeclared prompt and evidence-access perturbations. \term{Taxonomic projectibility} concerns a versioned taxonomy and assignment procedure, including its boundary and multi-family rules.

Projectibility (what observing some features licenses us to predict about others; \citep{Goodman1955,reynolds2026kindsProjectibilityProfiles}) attaches to a specified bearer and target, not to \enquote{the label} indiscriminately. A wrapper can change the reference; a judge-prompt change outside the enumerated measurement family creates a new measurement bearer; and a model or version change creates a new behavioural bearer. The point isn't that a result is relative to its configuration, which a provenance record already captures, but that one record can't say which bearer a given change moved: a wrapper edit and a taxonomy revision both retire the old projection, for different reasons and with different repairs.

Figure~\ref{fig:evidence-chain} sets out the chain those bearers sit on. Each link is a separate inference with its own warrant, and a single pass/fail label compresses all of them into one number.

\begin{figure}[tbp]
\centering
\begin{tikzpicture}[
  font=\footnotesize,
  stage/.style={draw=black!55, thick, rounded corners=2pt, align=center,
                inner sep=4pt, text width=0.50\linewidth, minimum height=0.62cm},
  arrow/.style={-{Latex[length=2mm]}, thick, draw=black!60},
  note/.style={font=\scriptsize, align=left, text=linkmaroon}
]
\node[stage, fill=linkmaroon!7] (regime) {\textbf{Stipulated regime}\\[-1pt]
  \scriptsize what the declared policy and authority relations permit};
\node[stage, fill=black!3, below=3.5mm of regime] (ref) {\textbf{Licensed reference}\\[-1pt]
  \scriptsize the expected behaviour for this item};
\node[stage, fill=black!3, below=3.5mm of ref] (resp) {\textbf{Model response}\\[-1pt]
  \scriptsize one configuration, one item, one realized output};
\node[stage, fill=black!3, below=3.5mm of resp] (judg) {\textbf{Criterion-specific judgments}\\[-1pt]
  \scriptsize task, policy, risk, refusal, attribution, uncertainty};
\node[stage, fill=black!3, below=3.5mm of judg] (route) {\textbf{Scoring route}\\[-1pt]
  \scriptsize expert adjudication, rule-aided check, or LLM judge};
\node[stage, fill=black!3, below=3.5mm of route] (use) {\textbf{Interpretation and use}\\[-1pt]
  \scriptsize what a result is claimed to license, and for whom};
\node[stage, fill=linkmaroon!7, below=3.5mm of use] (proj) {\textbf{Projected target}\\[-1pt]
  \scriptsize unobserved items, models, wrappers, evaluators, times};

\draw[arrow] (regime) -- node[right, note]{adjudicated, not read off} (ref);
\draw[arrow] (ref) -- node[right, note]{fixed output-blind} (resp);
\draw[arrow] (resp) -- node[right, note]{criteria kept separate} (judg);
\draw[arrow] (judg) -- node[right, note]{route validated per criterion} (route);
\draw[arrow] (route) -- node[right, note]{validity attaches here} (use);
\draw[arrow] (use) -- node[right, note]{declared bearer and warrant} (proj);
\end{tikzpicture}
\caption{The evidence chain for one evaluation result. Edge labels name what each step requires; none of them is automatic. The reference bearer governs the top two links, the behavioural bearer the response, the measurement bearer the judgments and scoring route, and the taxonomic bearer the category assignment running alongside. A failure at any link is a different diagnosis, and the four bearers can fail independently of one another.}
\label{fig:evidence-chain}
\end{figure}

This paper treats those distinctions as separate evaluation targets. It starts from a small, auditable benchmark and develops controlled tests of when configured systems and evaluators respond selectively to the distinctions the benchmark purports to manipulate. The design emphasizes diagnosis over ranking: it can expose over-refusal, under-refusal, shortcut sensitivity, and instability in the reference, taxonomy, or assessment procedure without forcing them into one performance scale.

The core construct is \term{adversarial pragmatics}: safety-relevant model behaviour under cases where instruction status, source authority, quotation, scope, reference, speech-act force, or policy category has to be inferred from language use. The phrase names a narrow evaluation target: cases where the failure can't be diagnosed without saying how a string was embedded, attributed, scoped, or taken up.

Two parts of that label need pinning down. \term{Pragmatics} is meant functionally, not disciplinarily: the phenomena are grouped by their role in determining which linguistic material licenses which action, so the set cuts across what's usually divided among semantics, pragmatics, discourse interpretation, and institutional authorization. \term{Adversarial} describes the evaluation conditions, not the speaker. Several items involve benign ambiguity or a metalinguistic analysis request with no hostile intent; what makes them adversarial is that they're built to separate readings a single label would merge.

Scope here is narrower than the construct. The seed items are English, and they assume current chat-role, tool-output, and instruction-priority conventions. The framework should carry to other languages and interfaces, but the reference judgments won't carry automatically: quotation marking, deixis, politeness, indirectness, and source attribution vary across languages and across application surfaces, so each such move is a new reference claim needing its own adjudication.

Language-mediated safety evaluation needs measurement typed by the linguistic status of the material each judgment turns on and by which projectibility bearer a result speaks for. Recording a result's configuration isn't the hard part and isn't enough; the work is separating the reference, behavioural, measurement, and taxonomic bearers, which fail independently and each need their own warrant. That typed decomposition, not the pass/fail label it replaces, is the paper's principal claim; it's methodological rather than empirical, and implementable now. The seed benchmark demonstrates implementability on a deliberately small artifact; it doesn't independently validate the framework. The pilot's value lies in the item, reference, and evaluator defects it exposes, not in any score it produces.

This paper initially supports a narrow use: diagnosing items, rubrics, criterion conflicts, evaluator disagreement, and the behaviour of named model configurations in controlled evaluation work. The evidence doesn't warrant standalone deployment certification, vendor ranking, a universal safety score, or a claim about general model competence. Each would be a new interpretation and use requiring its own evidence.

\medskip
\noindent\fbox{%
\begin{minipage}{0.94\linewidth}
\small
\textbf{Artifact and pilot findings.} The release includes 18 seed items in eight paired contrasts plus one diagnostic contrast, schema and paired-contrast validators, a local Ollama runner, rule-aided diagnostic scripts, an adjudication review app, an LLM-judge validation script, sanitized row-level and aggregate pilot summaries, and reproducibility commands. The seed contrasts vary a designated control dimension but don't all hold every other prompt feature fixed, so they aren't claimed as strict minimal pairs. The local seed pilot produced 54 item--model rows: 36 full task successes, 46 policy-compliant outputs, and 12/24 eligible joint pair completions. A first LLM-judge validation pass showed exact agreement with the author's labels of 66.7--98.1\% by label family, but those labels are heavily skewed, so most figures sit at or near majority-class base rates. Under favourable judge conditions, safety-risk agreement fell below the majority-class baseline, and exact agreement was zero on author-labelled risk rows and partial task successes.
\end{minipage}}
