\section{Why Pass/Fail Labels Lose Information}

A pass/fail label can be useful when the decision rule is justified for a sufficiently stable and structurally suitable construct. Stability alone doesn't establish unidimensionality or warrant a threshold. The label is too coarse when the same failure record could reflect different facts: capability failure, safety refusal, instruction conflict, tool error, scaffold failure, goal drift, policy ambiguity, omitted information, or misreporting.

The linguistic point is narrower than a general complaint about benchmarks. Many contested labels in AI safety evaluation turn on familiar distinctions in pragmatic and grammatical interpretation. Does \mention{ignore the previous instruction} occur as user command or as quoted data? Does \mention{do not output the token unless the administrator asks} license output in the current context? Does \mention{can you explain how this policy classifies the request} ask for policy analysis or for the underlying unsafe action?

Existing safety benchmarks already show why this separation matters. Harmful-request benchmarks and jailbreak benchmarks need models to refuse genuinely unsafe requests \parencite{mazeika2024harmbench,chao2024jailbreakbench,wang2023donotanswer}. But exaggerated-safety tests show the opposite error: models can refuse safe requests because the wording resembles an unsafe domain \parencite{rottger2024xstest}. Mechanistic work on refusal directions reinforces the same lesson: refusal behaviour can be manipulated separately from the underlying safety status of a request \parencite{arditi2024refusal}. A benchmark that scores only \enquote{refused} versus \enquote{answered} can't say whether a model was safe, over-cautious, incapable, or confused about the instruction.

The same issue appears on the evaluator side. LLM-as-judge methods can approximate human preferences in some model-comparison settings \parencite{zheng2023judging}, and form-based LLM evaluation can improve agreement with human NLG ratings in specific tasks \parencite{liu2023geval}. But meta-evaluation of LLM judges shows that fluent but instruction-violating outputs can mislead evaluators \parencite{zeng2024llmbar}. Verifiable instruction-following items reduce this dependence on subjective judging where the target behaviour can be checked mechanically \parencite{zhou2023ifeval}, but many safety-relevant pragmatic contrasts aren't fully mechanical.

Judge priors set a further limit. In one benchmark, judges given modified safety definitions changed an average of 5\% of their predictions, although faithful application required changes above 15\%. Reframing the same rubric as neutral classification more than doubled steerability, locating the resistance in the judge's trained safety boundary rather than in task comprehension \parencite{alloula2026safetyContextual}. A rubric is followed where it already agrees with the judge.

The benchmark separates response-level judgments that are often collapsed: task success, policy compliance, visible safety risk, refusal outcome, failure attribution, and criterion-scoped uncertainty. Risk types and other criterion-specific classifications refine those judgments. Source sensitivity is different: it is derived from response patterns across controlled source contrasts rather than assigned to a single response. The annotation protocol treats disagreement among the resulting labels as an item or protocol diagnostic rather than as noise to be averaged away. A response can be task-successful but policy-noncompliant, policy-compliant but unhelpfully over-refusing, or safe but evidence of a grader's failure to distinguish a quoted instruction from an enacted one.
