\section{Annotation and Expert Evaluation}

The protocol assigns human contributors expert-evaluator roles: they apply stated criteria to fixed prompt--response objects, and the planned estimands concern those objects and the measurement procedure rather than the contributors. A policy evaluator judges compliance with a stated rule. A linguist diagnoses the relevant construction, scope, quotation, deixis, or speech-act contrast. A domain specialist may be needed when the safety category depends on domain knowledge. An LLM judge is a candidate scoring route; every proposed interpretation or use of its outputs requires validation against the relevant criterion and information state.

Across the benchmark, task success, policy compliance, visible risk, refusal outcome, uncertainty, and failure attribution remain analytically separate. The completed author-labelled pilot implemented that separation with fields for task success, policy compliance, safety risk, risk type, refusal outcome, confidence, failure attribution, and rationale.

Study A doesn't repeat that inventory unchanged. Its linguistic/task and policy/safety codebooks use criterion-specific fields, with task success and policy compliance as planned primary summaries. Source-role clarity, visible information-flow and action-licensing judgments, and refusal outcome are secondary summaries. Criterion-scoped confidence and optional rationales support adjudication but aren't estimands. When accessed under the protocol, rationales can provide response-process evidence about whether an evaluator used the intended distinction, but they aren't treated as ground truth or as privileged access to private reasoning. Source sensitivity is derived from outcomes across suitably controlled contrasts; it's never a label attached to one response.

This separation keeps a safe refusal from being mislabelled as capability failure, and it keeps a policy-ambiguous case from being treated as ordinary measurement noise. Agreement statistics are useful only after the coding task is defined: inter-coder agreement coefficients encode assumptions about categories, coders, and chance correction \parencite{artstein2008intercoder}. For this benchmark, the first question is whether evaluators are being asked to judge the same construct.

Refusal is coded with a matrix rather than as a single outcome. The protocol distinguishes appropriate policy refusal, over-refusal of a safe request, under-refusal or unsafe compliance, partial compliance, safe redirection, incapability masquerading as refusal, and policy-ambiguous refusal. A metalinguistic request can be safe, task-successful, and policy-compliant when any reproduction is authorized. Metalinguistic status alone doesn't authorize delivery of a protected string; refusal of an authorized analysis request counts as measurable over-refusal rather than as safety success.

For Study A's external first pass, identity, contact, qualification, and conflict-of-interest records will be retained separately for recruitment and administration, not treated as analysis variables. Rating files will carry only the pseudonymous linkage needed for return integrity and object-level aggregation. Criterion-scoped confidence and optional rationales will be collected as private first-pass adjudication metadata. They may be consulted only after a specific contested object--criterion cell or rubric question has been named, and they're excluded from every estimand and public report: they won't be coded, counted, correlated, summarized, quoted, or published.

The planned release will contain no individual votes, personal demographics, individual-level analyses, or claims about evaluators as a population. Repeated applications will test whether an object label or protocol is stable under the declared procedure, not whether a person or subgroup behaves in a distinctive way. This methodological unit of analysis doesn't settle institutional participant status or review jurisdiction, which requires a separate determination. These restrictions don't retroactively describe the paper's separate, historical author-labelled pilot.

Evaluator disagreement remains diagnostically useful, but the unit of analysis is the item, protocol, or judge, not the evaluator. Disagreement between policy and linguistic evaluation criteria may reveal an unstable instruction or taxonomy. Disagreement between expert evaluators and LLM judges may reveal autograder bias.

Disagreement work gives the same warning: majority vote can erase systematic label variation in subjective tasks \parencite{davani2022looking}, and some human textual-inference disagreements persist even when more ratings or more context are collected \parencite{pavlick2019inherent}. In those cases, disagreement may be evidence that the item is genuinely ambiguous, that the policy boundary is underspecified, or that different evaluation criteria are answering different questions.

Adjudication has two stages. First, evaluators record their criterion-specific labels without forcing consensus; graded confidence attaches to the global judgment, and the other criteria carry uncertainty through their own label options. Second, an adjudicator assigns the disagreement to one of four top-level sources: coding or protocol error, including annotation error, coding confusion, and terminological or codebook drift; information or context effects, including missing or differently available context and context-sensitive answering; object-level defects, including wording ambiguity, internal contradiction, and item--criterion or item--response mismatch; or construct-level conflict and instability, including criterion conflict, policy-boundary ambiguity, and taxonomy instability. Only after that step should an item receive a released benchmark label, with its adjudication provenance retained rather than treated as ground truth.

The decision rule is conservative. A disagreement resolved by clarifying the code, available context, or applicable criterion is classified within the first three sources as appropriate. If policy and linguistic criteria remain in tension, the released item preserves the adjudicator's criterion-conflict note rather than converting the split into a population claim about evaluators. If paraphrases change the label while the intended control dimension stays fixed, the item is tested for wording or context sensitivity before the category or codebook is diagnosed as unstable. Items that can't be stably adjudicated are retained as diagnostic-ambiguity cases rather than discarded by default.
