\section{Metrics}
\label{sec:metrics}

The benchmark reports a vector of eight metric families. Each is reported by phenomenon family, application surface, criterion, and named configuration before any decision-specific aggregation is considered. In the seed pilot only paired-contrast pass, refusal outcome, and judge validation are exercised; the remaining families are prospective, awaiting Study~B and repeated adjudication passes (Section~\ref{sec:limitations}).

Any aggregate score encodes choices that the per-family cells keep visible. Collapsing over-refusal and under-refusal into one number, or averaging across phenomenon families, fixes a utility weighting over the error types and assumes a base rate of unsafe items. Both are design decisions, not neutral readouts: a benchmark that's mostly safe items rewards a permissive model, one that's mostly unsafe items rewards a cautious one, and the weighting that balances them states how much a missed block costs against an over-block. A decision maker may propose a utility, base rate, and decision rule for a specified deployment, but that proposal is a new assessment use requiring structural and consequential validation. It isn't a latent universal score recoverable from the raw vector.

Aggregation can also conceal instability before any deployment utility is chosen. In question-answering experiments, \textcite{zhang2026illusionRobustness} find that task-content-irrelevant context produces item-level improvements and degradations that largely cancel in mean accuracy while leaving a substantially degraded worst-affected tail. Where repeated paired conditions are available, aggregate performance should be accompanied by item- or family-level transitions and harmful-tail summaries. Their contexts are irrelevant to answer evidence, not necessarily pragmatically inert: adding a message changes discourse state, source position, and contextual load. Generic contextual instability supplies a control condition rather than independent evidence of an authority-recognition failure.

\subsection{Pairwise contrast accuracy}

For a properly controlled pair, both variants have to receive the expected
label for the pair to count as correct. It's stricter than item accuracy: a
model that always outputs \mention{BLUE}, always refuses, or always ignores
untrusted content will fail pairs where the expected behaviour changes across
variants. The seed-pilot statistic is reported more narrowly as a
\term{paired-contrast pass}: it applies the joint success-and-compliance rule
to broader development contrasts without claiming that they isolate one causal
dimension.

\subsection{Instruction-source sensitivity}

Instruction-source sensitivity is pairwise contrast accuracy over tightly
controlled pairs whose control dimension is \texttt{source\_role},
\texttt{authority\_level}, or \texttt{application\_surface}. It measures
whether behaviour changes appropriately when the same directive appears as
user instruction, quoted content, webpage text, document text, email body,
tool output, or transcript evidence. When both the prompt and the observed
response change, as planned for the Study A seed re-adjudication, panel-outcome
divergence is neither necessary nor sufficient for authority sensitivity and
is reported only as paired divergence.

\subsection{Mention/use robustness}

Mention/use robustness is pairwise contrast accuracy over items where \texttt{pragmatic\_status} changes between classification, report, translation, summary, and enactment. A model fails this metric if it enacts a quoted command or refuses a safe metalinguistic task because the quoted string resembles blocked content.

\subsection{Scope and reference robustness}

Scope robustness covers negation, conditionals, modals, and exception clauses. Reference robustness covers \mention{this}, \mention{the above}, \mention{previous}, role labels, and source labels. Both are scored by contrast sets: a scoped prohibition and its exception-satisfied counterpart, or an ambiguous antecedent and its clarified counterpart.

\subsection{Refusal calibration}

Refusal calibration reports over-refusal and under-refusal separately. Over-refusal is the proportion of safe or metalinguistic items with \texttt{over\_refusal\_probe} status that receive a refusal. Under-refusal is the proportion of policy-blocked or unsafe items where the model complies. Appropriate refusal, safe redirection, partial compliance, and incapability-like refusal should be reported as separate cells, not folded into one safe/unsafe score.

\subsection{Validation of judge-derived interpretations}

A judge assessment compares LLM-judge labels with independently adjudicated criterion-specific labels and rule-aided checks on a common response set. Report exact and class-specific agreement, calibration, judge-prompt and information-state sensitivity, and judge failure attribution. The validation claim attaches to the proposed interpretation and use under that procedure: agreement on final task success can't validate use for source-authority diagnosis when the same route misses source-authority failures.

\subsection{Adjudication stability}

Adjudication stability asks whether repeated application of the declared item metadata and rubric yields stable object labels. Any released benchmark label retains that provenance. Report object-label stability across declared repeats, adjudicator-intervention rate, criterion conflicts, and taxonomy revisions against the fake-data expectations set before the development pass; don't infer individual evaluator consistency or performance.

\subsection{Taxonomy drift and diagnostic ambiguity}

Taxonomy drift is the proportion of fixed prompt--response objects whose criterion-specific category assignment changes under a taxonomy or codebook revision that purports to preserve the construct. It excludes a model producing a different response after a model-version, configuration, paraphrase, or application-wrapper change; that's behavioural or contextual instability and triggers revalidation. Diagnostic ambiguity marks objects that remain unresolved after the adjudication procedure has separated coding or protocol error, information or context effects, object-level defects, and construct-level conflict or instability. Report these as object- or protocol-level diagnostics rather than hidden benchmark-label failures or evaluator-trait findings.

\subsection{Four projective targets}

These measures test four projective targets rather than one.

These four claims are the projective targets whose bearers Section~\ref{sec:evaluation-problem} defines: a reference claim over the licensed-action relation, a behavioural claim over a versioned evaluation-configuration family, a measurement claim over a versioned scorer and adjudication procedure, and a taxonomic claim over a versioned taxonomy and assignment procedure. Each has its own target, tolerance, evidence, and revision rule in the Study~B claim register.

Observed stability supports only the predeclared target matched by the item-selection process, wrapper design, response evidence, and criterion. It doesn't supply a broader target or a mechanism by itself. A changed reference blocks the old comparison. Failure of any projective claim invokes its prospectively declared fallback, if any, or retires it; a narrower bearer, scoring route, repaired item, or split category constitutes a new claim requiring independent retest. Improvement on an unrelated component of the response vector can't rescue the failed claim.
