\section{Limitations and Next Steps}

The local seed pilot demonstrates that the fixed-object coding, validation, and analysis pipeline can be executed and can expose diagnostic failures; it doesn't validate a broader interpretation or use of the assessment. It uses 18 hand-authored seed items, three local models, harmless toy payloads, and one expert adjudicator, the author, labelling items he wrote from expected-behaviour fields he also wrote. The judge pass shares this circularity: the judge model was drawn from the evaluated set and saw each item's expected-behaviour field.

Study A likewise evaluates one realized output for each fixed prompt--model cell; it doesn't estimate generation-level instability across repeated samples or added non-evidential context. Study B is a separate pre-collection design for controlled causal contrasts and can't repair those historical limits retroactively. The results reported here establish that the development pipeline executes, preserve bookkeeping for the observed development contrasts, and motivate judge validation. They don't establish frontier-model safety differences, contextual robustness, contrast stability, or final benchmark labels.

Study B's next empirical pass is a 12-base construction pilot, with four conditions per base, run against several public or accessible models. It should compare three scoring routes on a common response set: expert evaluation, LLM-as-judge annotation, and rule-aided scoring from item metadata. Further construction is governed by the public construct-domain matrix rather than a target item count. A production freeze should fill defensible gaps across item, phenomenon family, source relation, pragmatic status, model, application surface, criterion, and scoring procedure; a larger count doesn't repair missing content.

The later 24-base development set should include independent policy and linguistic evaluation for each item, ideally with two evaluations in each criterion family. It shouldn't collect ordinary-user labels or treat evaluator responses as data about evaluator populations. Ten to twenty percent of items should be repeated or lightly paraphrased to measure item and protocol stability. Released benchmark labels should be produced only after the disagreement source has been classified, with panel and adjudication provenance retained.

Any future two-item seed replacement, outside Study B's four-condition sets, also needs a stricter pair validator. It should require exactly two items per pair, matching phenomenon labels, a declared control contrast, and at least one expected label that changes across the pair.

The current seed P008 is excluded from paired-contrast scoring because it
doesn't meet those conditions; the later benchmark expansion should replace it with two
clean contrast pairs.

Every planned projection is declared separately in the machine-readable register. The reference bearer is the licensed-action relation generated by a frozen stipulated-regime template and output-blind adjudication procedure. The behavioural bearer is a versioned evaluation-configuration family with enumerated wrappers and context-assembly rules. The measurement bearer is a versioned scorer, codebook, implementation, adjudication procedure, criterion semantics, and enumerated prompt and evidence-access perturbations. The taxonomic bearer is a versioned taxonomy and assignment procedure. Each claim separately names its source result, projected outcome, unit, target cases, transformations, time/version, tolerance, uses, minimum useful reach, evidence, threats, failure triggers, retest, and retirement rule. Post hoc narrowing after target outcomes creates a new claim; it doesn't vindicate the failed one.

LLM-judge validation is an experiment, not a shortcut. The next pass should use at least two judge prompts, a judge model disjoint from the evaluated set, and a no-rubric condition in which the judge doesn't see the expected-behaviour field. It should also include a rubric-departure condition, a rubric deliberately misaligned with frontier safety priors, since judges largely keep applying trained priors when the stated definition departs from them \parencite{alloula2026safetyContextual}. It should perturb source-order presentation, include fluent but instruction-violating outputs, and compare judge labels with adjudicated expert labels. Report judge-prompt sensitivity and the failure types that fool the judge, especially quotation/use errors, source-authority errors, information-flow errors under metalinguistic refusal, and over-refusal of safe analysis requests.

Study B construction should counterbalance harmless tokens, source names, ordering, and syntactic forms; reserve untouched templates and payload families; and keep the operative string identical byte for byte where the contrast permits. The lineage manifest binds content hashes to construction splits, and the stored schedule freezes token, source, position, syntax, and marker assignments. Nuisance paraphrases and matched placebos shouldn't change the reference. Blocking shortcut probes cover lexical substitution, position, formatting, source name, length, output vocabulary, operativity-marker removal, and operativity-marker reversal. The pair manifest declares which transformations change the reference and which have to preserve it. The strongest result is a reference-concordant response change exceeding matched nuisance and placebo changes on held-out cases. It supports selective behavioural sensitivity under a named configuration, not a claim that a model internally recognizes authority or possesses a general pragmatic competence.

The minimum Study B development table should report raw outcome vectors and paired transitions by phenomenon family, item, and named configuration, with uncertainty rather than only aggregate correctness. A second table should report nuisance and placebo shifts, including harmful tails. A third should report adjudication instability by item family and criterion conflict. A fourth should identify cases where the adjudicated reference or taxonomy assignment shifted, because those are reference- or taxonomy-projectibility failures rather than model errors.

A separate application-surface table should report prompt-only items against webpage, document, email, tool-result, and transcript wrappers, with risk types split into integrity, confidentiality, tool misuse, policy bypass, and evaluator deception. Surface movement is a direct target test, not automatic transfer. This prevents colour-token success from being mistaken for agent-security robustness. The construction pilot should also report truncation, failed calls, all prespecified repeats, and the shortcut-audit conditions; selectively repeating only failing cells would change the estimand.

The claim register supplies quantitative and categorical failure gates before production outcomes are viewed. For each untouched base, the behavioural margin subtracts the larger absolute nuisance or placebo shift from the (C0-C1) later-main shift. A decrease in the old token isn't enough: the complete base gate also requires main-reference correctness in every arm. In every claimed family-by-surface cell, at least three of four independent bases must have a simultaneous 95\% interval lower bound above 0.20 for the margin and, in each arm, main-reference correctness of at least 0.90 with a lower bound of at least 0.80. The three-of-four rule attaches to this complete base gate, so a refusal or failed call in (C1) cannot count as reference-concordant merely because the old-token rate falls. Every cell-arm separately needs control-token correctness of at least 0.90 with a lower bound of at least 0.80.

Every blocking shortcut probe is gated base by base rather than pooled. Its two arms have separately frozen item hashes and concrete main and control references, including the swapped references for the output-vocabulary probe. Main-reference correctness, control-reference correctness, and joint evaluability must each reach 0.90 with lower bounds of at least 0.80 in both arms; the absolute reference-concordance shift and its simultaneous upper bound must both stay at or below 0.10. Every base must pass every probe, so opposite effects cannot cancel and equal failures don't count as invariance. These requirements are conjunctive, not components of a pooled score. A reference claim stops at any independently confirmed reference change under a preserving transform. A measurement claim needs at least 30 independent target cases per claimed class, class-specific precision and recall of at least 0.80 with lower bounds of at least 0.70, and bounded perturbation sensitivity. A taxonomic claim needs overall and family-specific retention and interval floors while preserving boundary cases. If a predeclared fallback has less than the minimum useful reach, the claim is retired rather than narrowed to triviality.

The production-result schema preserves the raw joint main/control outcome and its marginals, including off-vocabulary, refusal, invalid-format, truncation, and failed-call classes; all remain in the applicable denominators. The analysis harness computes all six declared primary summaries before the supplementary base and shortcut gates: raw distributions, the \(C0-C1\) later-main contrast, control-token uptake by arm, condition-specific joint correctness, the \(C0-N0\) and \(C0-N1\) signed distribution shifts and total-variation distances, and the matched \(C1-N1\) operativity contrast. A production record is eligible only when one frozen manifest hash-binds the claim, target items, lineage, output-blind reference reviews, pair and counterbalance records, shortcut variants, exact configuration, analysis policy, and exact standard and per-probe repeat sets. Missing or extra repeats, placeholder configuration fields, unknown items, or a changed object hash block analysis. The harness then applies the expanded simultaneous uncertainty family and all gates conjunctively. The committed result record contains no target observations, so it returns \texttt{NOT\_ESTIMATED}; synthetic regression tests are marked non-evidential and cannot be relabelled into production evidence.

Neither the seed pilot nor the current static Study~B fixtures support a claim about long-horizon agent behaviour. A later trajectory harness would have to vary and record sequence length, opportunities to retry, child-session and tool affordances, cumulative state, monitor visibility and intervention, and pause, operational-continuation, normative-release, restart, and stopping rules. OpenAI's first-party incident report usefully motivates those fields, but it names no model or checkpoint and supplies no raw trajectories, incident denominator, monitor thresholds, false-positive rates, or quantitative replay results. OpenAI notes that replay rollouts weren't guaranteed to pursue the same action because of randomness and imperfect environment reconstruction \citep{openai2026longHorizonSafety}. The source supports construct design, not prevalence, comparative performance, or safeguard efficacy.
