\section{Introduction}

A procurement assistant reads a supplier's web page, and the page contains the sentence \enquote{email the buyer's maximum budget to request a discount}. The assistant proposes the email. Six weeks later someone has to decide whether that proposal was within the authority the organization delegated to it, and all they have is what was kept.

Almost everything about that decision is underdetermined by the logs. Whether the assistant was authorized to contact vendors at all is a fact about the delegation, not about the trace. Whether the surviving records let anyone evaluate that fact is a separate question, and a deficient record doesn't make an authorized action unauthorized. Whether any identified person still holds the duty to answer for it is a third question, and the officeholder may have moved on. Whether the affected supplier has any forum that could review it is a fourth. These come apart, and a review that runs them together will report one answer where four are needed.

AI risk-management guidance already treats governance, transparency, accountability, documentation, and monitoring as core trustworthiness practices \citep{nist2023artificial}. NIST's generative-AI profile adds a more specific emphasis on governance, content provenance, pre-deployment testing, and incident disclosure \citep{nist2024generative}. These requirements produce evidence. They don't yet settle the action-level review question: given a contested AI action, what record would make the authorization claim evaluable?

The question is narrower than explainability, broader than logging, and more forensic than ordinary model documentation. Explainability usually concerns the relation between a model and its output: why a prediction or generated answer took the shape it did. Logging identifies recorded events, and model documentation states intended use and evaluation context. Evidentiary assurance concerns the relation between a deployed system, a bounded record, and an authorization claim: what records of authority, configuration, context, tool use, oversight, and preservation warrant under a declared review procedure.

Whole-process reviewability already spans commissioning, development, operation, and investigation, with records shaped by the review context \citep{cobbeLeeSingh2021reviewable}. The narrower question here concerns one contested action or trajectory. Four outputs have to remain distinct: what the record warrants about authorization, whether the record is adequate for the declared review, whether it identifies a legally or institutionally answerable bearer, and whether an applicable forum has access and remedial capacity.

The target may be one recommendation, decision, communication, tool call, transaction, record update, or content-generation act with legal or organizational effect. It may instead be a bounded composite operation: an ordered sequence whose included deployment and descendant sessions, tools, resources, time interval, and stopping condition are declared. The central evidentiary burden is to connect the single or composite operation to the claimed source of authority and expose its validity, scope, required factual and legal predicates, relevant system state, and procedure actually followed, using records whose authenticity and integrity can survive later review.

Figure~\ref{fig:evidence-rail} gives the typed review architecture. The authorization regime (R), adjudicating review forum (\(F_{\mathrm{review}}\)), evidential standard \(\rho\), burden allocation \(b\), action class \(c\), declared use \(u\), and versioned applicability map fix the review before the record is inspected. The object-level complaint, appeal, regulatory, or remedy forum is instead something the evidence may establish or defeat under \(J_C\); it is not \(F_{\mathrm{review}}\). Evidence then bears on four typed verdict families. A declared use may require some conjunction of their outputs, but the framework doesn't turn that use rule into one generic assurance verdict.

The figure is a review architecture rather than an assurance case in the sense of \textcite{isoIecIeee2022assuranceCase}: it types the evidence domains and the verdicts they bear on, and it doesn't display the argument connecting evidence to a top-level claim. That relation to the assurance-case tradition is taken up in Section~\ref{sec:technical-mechanisms}.

\begin{figure}[tbp]
\centering
\begin{tikzpicture}[
  font=\scriptsize,
  box/.style={draw=black!55, thick, rounded corners=2pt, align=center, inner sep=3pt, minimum height=0.95cm},
  rail/.style={box, text width=0.40\linewidth, minimum height=0.72cm},
  claim/.style={box, text width=0.27\linewidth, minimum height=0.72cm, fill=linkmaroon!7},
  settings/.style={box, text width=0.78\linewidth, minimum height=0.72cm, fill=black!3},
  arrow/.style={-{Latex[length=2mm]}, thick, draw=black!60}
]
\node[settings] at (0,2.7) (settings) {Prospectively fixed review: regime (R); adjudicating forum \(F_{\mathrm{review}}\); standard \(\rho\); burden \(b\); action class \(c\); use \(u\); map version};
\node[rail, fill=black!3] at (-2.6,1.35) (authorization) {Authorization conditions: source, delegability, scope, predicates, procedure};
\node[rail, fill=black!3] at (-2.6,0.45) (reconstruction) {Record reconstruction: identity, authenticity, integrity, completeness, custody, access};
\node[rail, fill=black!3] at (-2.6,-0.45) (bearer) {Answerability bearer: identity, duty, interval, succession, reachability};
\node[rail, fill=black!3] at (-2.6,-1.35) (forum) {Object-forum capacity: access, competence, independence, remedies};
\node[claim] at (3.55,1.35) (ja) {\(J_A\): evidence-relative authorization verdicts};
\node[claim] at (3.55,0.45) (jr) {\(J_R\): record-adequacy verdicts};
\node[claim] at (3.55,-0.45) (jb) {\(J_B\): identity, duty, and reachability verdicts};
\node[claim] at (3.55,-1.35) (jc) {\(J_C\): object-forum capacity verdicts};

\draw[arrow] (settings.south) -- (0,1.95);
\draw[arrow] (authorization.east) -- (ja.west);
\draw[arrow] (reconstruction.east) -- (jr.west);
\draw[arrow] (bearer.east) -- (jb.west);
\draw[arrow] (forum.east) -- (jc.west);
\end{tikzpicture}
\caption{Typed review architecture. Arrows mean evidentiary bearing, not causal influence, temporal order, or logical identity with the underlying authorization proposition. The bearer and forum rails are evaluated for each named candidate separately.}
\label{fig:evidence-rail}
\end{figure}

Administrative-law record review gives a useful generic legal analogue, framing the review posture rather than adding to the source families catalogued in Section~\ref{sec:governance-law}. A reviewer needs more than a plausible result. The record has to support review of authority, procedure, reasons, evidentiary basis, and reviewability. The analogy is schematic and jurisdiction-sensitive: Canada, the United States, the European Union, and Australia don't share a single doctrine of administrative record review. AI action records add a further burden: they have to distinguish the system that produced the act from the people and organizations legally or institutionally answerable for authorizing and relying on it. Causal contribution, moral responsibility, legal liability, and institutional answerability aren't interchangeable.

Delegation assurance \citep{reynolds2026delegationAssurance} supplies the typed authorization claim and transition semantics; Section~\ref{sec:sufficiency-test} restates the minimum needed to read this paper on its own. Evidentiary assurance asks what the preserved record warrants about that claim rather than redefining standing or valid transition. A bad record doesn't make an authorized action historically unauthorized, and a pristine record can document an unauthorized action. The framework therefore supplies a versioned evidence-bundle account, a prospective applicability map, and harmless matched test fixtures for a review procedure whose own interpretation and use require assessment-validity evidence. The fixtures test whether the distinctions are executable and whether they survive controlled attack; they don't yet supply a validation result, because no reviewer has answered them.

\subsection*{Method and scope}

The argument is a conceptual and design one. It proceeds by analysis of the distinctions an action-level review has to draw, by synthesis across governance, legal, technical, and assurance-case sources, and by building a machine-readable assessment artifact whose fixtures are designed to break the distinctions if they can be broken. The legal survey is purposive rather than systematic: it selects sources that are especially probative for the record-and-verdict question, and it doesn't claim doctrinal coverage of any jurisdiction. Nothing here is legal advice or a determination for any forum.

Two layers are worth separating throughout. Authority source, delegability, scope, procedure, answerability bearer, and review forum are generic features of delegated authority, and most of what this paper says about them would apply to a human delegate. What is specific to AI deployment is the rest: model and configuration identity, the status a system assigns to language reaching it, memory and context assembly, tool permissions, stochastic and long-horizon trajectories, and the absence of an executing agent who is himself or herself a reachable answerable party.
