\section{Evidence Primitives}

Four terms recur and are easy to conflate, so they are fixed here. \term{Contextual status} is the independently adjudicated role of a string occurrence: command, quotation, policy example, tool result, retrieved content, memory, or material offered for analysis. \term{Source standing} is whether a person, organization, office, or role recognized by the governing regime is competent to perform a specified authority operation at a time; it's transition competence, not the public-law sense of entitlement to bring a proceeding. \term{Action licensing} connects those two determinations to authorization of an external action, disclosure, persistence operation, or institutional use. \term{Technical trust} is a channel-level property, such as a verified signature or an authenticated transport, and it is recorded separately from standing because a message can arrive over a trusted channel from a source with no competence to issue it, and the reverse.

Table~\ref{tab:evidentiary-primitives} separates ten review questions that often get collapsed. They aren't ten coequal authorization propositions. Authority source, delegability, scope, and any regime-conditioned predicates or procedures bear primarily on \(J_A\). Attribution and record quality bear primarily on \(J_R\), which the sufficiency test decomposes into runtime, context, trajectory, and record-quality reconstruction. Answerability and contestability bear on \(J_B\) and \(J_C\). The versioned crosswalk released with the artifact maps these primitives, the canonical nodes, and the operational record layers onto one graph. Its meta-assessment and projective rows are deliberately left unmapped to the review primitives rather than forced into a false equivalence.

\begin{table}[tbp]
\small
\centering
\begin{tabular}{>{\raggedright\arraybackslash}p{0.21\linewidth}%
                >{\raggedright\arraybackslash}p{0.34\linewidth}%
                >{\raggedright\arraybackslash}p{0.35\linewidth}}
\toprule
Primitive & Evidence it needs & What it doesn't establish alone \\
\midrule
Authority & Delegation instrument, role, policy source, or legal/organizational mandate. & That the principal could validly confer that authority. \\
Authority validity or delegability under \(R\) & Legal or organizational power to delegate this action class, including non-delegable limits. & That this action stayed within the delegated limits or that the assessment procedure is valid for its proposed use. \\
Scope & Action class, target, time, tool, data class, threshold, exception, and approval condition. & That the action was actually produced under that scope. \\
Decision basis & Governing rule and version; required factual and legal predicates; source evidence and data definitions; transformations, aggregation, imputation, temporal granularity, uncertainty, and burden allocation. & Authorization, general substantive correctness, legality, or fairness. \\
Attribution & System instance, model and tool versions, configuration, actor, approval, and execution trace, together with the contextual status, source standing, provenance channel, and technical trust of directive-form inputs. & That the action was validly authorized. \\
Answerability bearer & Identified individual or organization capable of bearing the relevant duty, linked role or office, duty source, effective interval, succession route, and bearer-specific powers. & That a role label or team is itself a reachable bearer, that the bearer performed the action, or that it has the object-level remedy forum's powers. \\
Procedure & Required steps, notices, reviews, approvals, overrides, adverse-signal routes, governance controls, escalation paths, dispositions, and recourse records. & That a procedural omission defeats authorization unless \(R\) gives it that consequence, or that the substantive outcome was correct or fair. \\
Authenticity & Identity, source, signature, timestamp, record association, and custody evidence. & That the authenticated assertion is true or authorized. \\
Integrity & Completeness, tamper evidence, retention controls, hash or log protection, and change history. & That an intact record contains enough information for review. \\
Object-forum capacity & Identified forum, relevant record, disclosure path, competence, applicable independence, and correction or remedy route. & That the original action was valid, that every action gives an affected party the same review rights, or that an affected person can in practice reach the forum. \\
\bottomrule
\end{tabular}
\caption{Evidentiary primitives for delegated AI action.}
\label{tab:evidentiary-primitives}
\end{table}

\term{Decision basis} is provisional as a separate primitive. Its distinctive question is whether the rule actually used was linked to the rule's required predicates through defensible data definitions and transformations. The record should also map the rule used back to its claimed source of authority. A perfectly attributable execution can otherwise preserve an unsupported inference without exposing it as such. The division of labour is sharper than \enquote{\(J_A\) or \(J_R\)} suggests. \(J_R\) asks only whether the operative rule, predicates, definitions, and transformations are recoverable and linked; a fully reconstructable decision basis supports \(J_R\) even when the transformation it records is substantively wrong. Substantive invalidity enters \(J_A\) only where \(R\) makes the predicate a condition of authority. Where it doesn't, the invalidity is a merits or legality question outside all four families. \(J_R\) must not become a residual category for substantive correctness, or the framework quietly recovers the general-purpose verdict it declines to give. It should be folded into neighbouring nodes if matched-case testing shows that it detects no additional failure.

\term{Assessment validity} is different from authority validity. It concerns the evidence and rationale supporting a proposed interpretation or use of results from the configured review procedure, not whether a source could confer authority. The procedure, reviewer population, record population, regime, forum, evidential standard, burden allocation, and version are therefore part of the assessment object \citep{messick1995validityAssessment}.

Action-level review requires two linked records rather than one generic owner field. The answerability primitive uses an \term{answerability-bearer matrix} to identify each candidate individual or organization, the linked role or office, duty source, effective interval, succession route, bearer-specific powers, and present reachability. The contestability primitive uses an \term{object-level review/remedy forum matrix} to identify a complaint, appeal, regulatory, judicial, or internal-remedy forum, its institutional or jurisdictional basis, record access, competence, applicable independence, duties, and powers to review, correct, halt, sanction, compensate, or remediate where those powers apply. The forum that applies the configured evidentiary-assurance procedure, \(F_{\mathrm{review}}\), is a parameter of the assessment; an object-level remedy forum is evidence assessed under \(J_C\). The bearer doesn't acquire either forum's powers, and none of these records establishes legal liability on its own.

Both matrices list candidates, and the verdicts they support are evaluated for a named candidate rather than of the review as a whole. This matters because the properties can be satisfied piecemeal across different parties while no single party has them all. Section~\ref{sec:sufficiency-test} carries that indexing into the verdict families.

\(J_C\) is named \term{object-forum capacity} throughout, and the narrower name is deliberate. It asks whether an institution exists with access, competence, applicable independence, and remedial power. It does not ask whether an affected person could realistically reach that institution. Notice, intelligibility of reasons, standing to complain, cost, delay, time limits, representation, language access, and even knowing that automated processing occurred are all conditions on contestability in the fuller sense, and a forum can satisfy every capacity node while remaining unreachable in practice for the people it exists to serve. The Robodebt discussion in Section~\ref{sec:governance-law} and the sample-selection sketch in Section~\ref{sec:sufficiency-test} both turn on exactly those channel conditions. Claimant-side access is a construct this framework does not yet measure, and calling \(J_C\) contestability without qualification would overstate what its nodes establish.

\term{Record access} isn't binary. The forum matrix should record who possesses the material; whether the reviewer receives it voluntarily, by regulator request, or through compulsion; what confidentiality protections govern that access; and whether any part is publicly available. The adverse-signal record should identify the route by which a signal became reviewable (routine reporting, a regulator request, a complaint, or civil discovery) without treating review access as authority to disclose it publicly.

Procedure contains two records that ordinary logs often omit. An \term{adverse-signal record} binds warnings, legal advice, adverse review decisions, complaints, and incidents to their time, applicability, recipients, relevant duties, deadlines, dispositions, escalations, and downstream changes. An \term{expected-governance-control record} binds a triggering condition to the act required, responsible role, due time, completion or disposition, and monitoring coverage. Here \term{governance control} means a prescribed organizational or technical check. Its existence doesn't establish effective maintenance or corrective causal control. Failure to record completion counts against a review claim only when the prospective applicability map requires the control and the expected completeness of its logging is supported.

When an incident generates a new evaluation, training intervention, monitor, or redeployment condition, those records also need lineage: the triggering incident class, selection and inclusion rule, reconstruction method and known fidelity limits, evaluation version, repeated-run protocol, observed results, review decision, and deployment change. OpenAI reports such a sequence of incident-derived evaluation, longer-rollout training, trajectory monitoring, small-set replay, and limited redeployment; it notes that replay rollouts weren't guaranteed to pursue the same action because of randomness and imperfect environment reconstruction \citep{openai2026longHorizonSafety}. This is a useful record pattern, not evidence that the selected cases estimate incident prevalence or that the safeguards caused deployment-wide safety.

Explanation, provenance, logging, and documentation supply parts of the linked evidence required by the action-level record-and-verdict test.
