\section{Governance and Legal Sources}

The source base is cross-jurisdictional, with emphasis on Canada, the European Union, the United States, and Australia where official sources are especially probative. The relevant sources fall into seven overlapping families.

Agency law supplies one useful vocabulary for the relation: a principal authorizes an agent to act on the principal's behalf, and actual authority may be express or implied \parencite{cornellWexAgency,cornellWexActualAuthority}. This is an analogy, not a claim that every statutory delegation, organizational authorization, subdelegation, or automated support arrangement creates an agency relationship. In ordinary agency, the agent can be a responsible legal or institutional actor. In AI deployments, executing an action doesn't itself supply an independently reachable answerability bearer; the record has to identify candidate answerability bearers among people and organizations, even when amendment and override powers are distributed across a delegation chain \parencite{cornellWexPrincipal}.

Activity-based agent governance supplies an external enforcement layer. \textcite{rajagopalan2026governingAgenticAI} argues that bare legal personality is neither necessary nor sufficient for identification, recovery, deterrence, or cessation and proposes registration, identification, verification, financial responsibility, lifecycle traceability, and suspension. Those layers address who can be reached, made financially responsible, and stopped. They don't establish that a particular action was validly authorized. This paper uses only that bounded enforcement claim and takes no position on personhood's other moral, recognitional, claimant-side, or expressive functions.

AI-governance instruments require logs, documentation, intended-purpose limits, instructions for use, human oversight, impact assessment, monitoring, recourse, and disclosure. The EU AI Act requires high-risk systems to support traceability through logging capabilities, requires instructions for use, and assigns deployers obligations around use, input data, monitoring, serious-incident reporting, and human oversight by people with competence, training, and authority \parencite{europeanUnion2024aiAct}.

Canada's Algorithmic Impact Assessment asks comparable operational questions about project authority, system accountability, audit trails, system-produced reasons, recourse, human override, documentary evidence, and changes in system functionality or scope \parencite{canadaAIA2026}. These instruments generate evidence, but they usually stop short of specifying what evidence is sufficient to support a finding that a concrete action stayed within delegated authority.

Proposed U.S. automated-vehicle legislation supplies a sectoral illustration. The introduced SELF DRIVE Act of 2026, H.R.~7390, would require a manufacturer-prepared safety case for each automated driving system (ADS) version. The case would address unreasonable crash, death, or injury risk. Required content includes a complete description of the \term{operational design domain} (ODD), performance of the dynamic driving task within that domain, hazard-analysis methods, acceptance criteria, crash response, and system competencies. The manufacturer would provide the case to the Secretary on request. These are proposed requirements, not current law, and the safety case concerns a version-level system-safety claim rather than an action-level authorization verdict \parencite{usCongress2026selfDriveAct}.

The bill separately proposes a repository for defined covered crashes and aggregate ADS-engaged mileage, with access for relevant state transportation regulators and public release subject to protections for confidential business information. The introduced text doesn't require public release of the safety case itself. Nor do its listed minimum crash fields alone identify the operative ADS version, whether the vehicle remained within its ODD, runtime configuration, authority path, or the completion or disposition of any triggered control. If enacted, an implementing rule could require more. A safety case, an incident-data repository, and the evidence needed to reconstruct a particular action are three distinct objects \parencite{usCongress2026selfDriveAct}.

Hinkle's congressional testimony makes the access question explicit. He characterizes automated driving as a manufacturer's domain-bounded promise and uses the General Motors ignition-switch litigation to argue that discovery can expose internal defect knowledge and decisions not to implement proposed fixes. He contrasts that process with confidential arbitration lacking comparable public records, ordinary evidentiary access, and meaningful appeal. This is advocacy testimony, not an adjudication or a general rule that all safety material must be public. For evidentiary assurance, the narrower implication is that possession by the record-holder, disclosure to a regulator or reviewer, legally compelled access, confidentiality protections, and public availability are distinct fields \parencite{hinkle2020autonomousVehiclesTestimony}.

Financial-reporting law supplies a useful institutional analogue. Sarbanes--Oxley Section 404 requires internal-control reporting that states management's responsibility for adequate controls and assesses their effectiveness; the SEC's implementing rules require management to identify the framework used for that evaluation, and Exchange Act Rule 13a-15 requires a suitable, recognized control framework for evaluating internal control over financial reporting \parencite{sarbanesOxley404,sec2003managementReport,secExchangeRule13a15}.

SOX concerns financial reporting, not machine action, but its structure is useful. Investors aren't asked to trust a company's bare assertion that the numbers came out right; management has to maintain a reviewable control environment around the reporting process. Evidentiary assurance asks for an analogous control record tying AI action to authority, scope, procedure, integrity, and accountable institutional actors.

Electronic-evidence and electronic-signature law already has mature concepts for authenticity, integrity, attribution, secure signatures, and tamper detection. The Canada Evidence Act places the burden of authenticating an electronic document on the party seeking admission and ties the best-evidence rule to proof of the integrity of the electronic-document system or an applicable evidentiary presumption \parencite{canadaEvidenceAct1985}. PIPEDA Part 2 and Canadian e-signature guidance add related primitives: signer identity, signer authority, association between signature and record, record retention, timestamps, transaction records, and integrity of the signed data and supporting information \parencite{pipeda2000,canadaESignatures2025}. Those doctrines aren't AI-specific, but they provide strong analogues for AI action traces.

The analogy is limited. Authenticity helps establish that a record matches its proponent's description; integrity helps establish that the record hasn't been altered in an impermissible way; weight concerns how persuasive the record is once considered. None of those conclusions is the same as sufficiency for delegated authority. An authentic and intact trace can still document an overbroad, invalid, or misapplied delegation.

Public failures show the practical stakes. The Robodebt Royal Commission found that employer-reported income over an annual or other extended period was allocated evenly across fortnights even though the governing assessment depended on actual fortnightly income; where a recipient didn't provide accepted alternative information, that transformation could become the basis for a debt. This was a \term{decision-basis} failure: a reconstructable calculation didn't supply the factual predicate the governing rule required \parencite{robodebtRoyalCommission2023}.

The report also documents failures in the circulation, disposition, or practical consequence of contrary legal and policy advice, adverse tribunal decisions, complaints, and warnings. And channel access affected which objections became visible while nonresponse helped trigger averaging. A jurisdiction-bounded rule-of-law analysis likewise emphasizes the annual-to-fortnightly inference, the corrective burden placed on affected people, and barriers of notice and review \parencite{zalnieriuteBennettMosesWilliams2019ruleOfLaw}. Predicate and transformation, adverse-signal routing, and observation selection are three distinct failures. They aren't one generic absence of logs \parencite{robodebtRoyalCommission2023}.

The Commission recommended clear review paths for affected people, plain-language notice that automated decision-making is used, and public availability of business rules and algorithms for independent expert scrutiny \parencite{robodebtRoyalCommission2023}. Its findings don't imply that every warning defeats an action or that improved records would have made the scheme lawful. They show why an assurance record needs warranted decision-basis lineage, expected-control evidence, and warning routes attached to identified duties and corrective powers. The governing forum determines the legal significance of any particular item.

Table~\ref{tab:legal-governance-primitives} maps these source families to their evidentiary contributions and remaining gaps.

\begin{table}[tbp]
\small
\centering
\begin{tabular}{>{\raggedright\arraybackslash}p{0.22\linewidth}%
                >{\raggedright\arraybackslash}p{0.28\linewidth}%
                >{\raggedright\arraybackslash}p{0.40\linewidth}}
\toprule
Source family & Evidentiary contribution & Remaining gap \\
\midrule
Agency law & Authority, validity, scope, attribution, accountability. & AI execution doesn't itself supply an independently reachable answerability bearer. \\
Activity-based agent governance & Registration, identification, verification, financial responsibility, lifecycle traceability, and suspension. & Reachability, recovery, and cessation don't establish that the particular action was authorized. \\
AI and sectoral safety governance & Safety cases, operating-domain scope, monitoring, incident reporting, denominators, oversight, recourse, and traceability. & Version-level safety cases and incident summaries don't necessarily make a particular action reconstructable or contestable. \\
Financial-reporting controls & Management responsibility, control frameworks, effectiveness assessment, attestation, change monitoring. & Internal control over reporting isn't yet an action-level standard for delegated AI systems. \\
Electronic evidence & Authenticity, integrity, attribution, preservation, evidentiary presumptions. & Document integrity doesn't by itself show that an AI action was authorized. \\
Electronic signatures & Identity, authority to sign, record association, timestamping, transaction evidence. & Signature validity doesn't generalize automatically to multi-step agentic action. \\
Public automation failures & Decision basis, adverse-signal routing, accessible channels, review paths, business-rule disclosure, independent scrutiny. & Post-failure findings still have to be translated into deployable evidence bundles and forum-specific legal tests. \\
\bottomrule
\end{tabular}
\caption{Evidentiary contributions from legal and governance sources.}
\label{tab:legal-governance-primitives}
\end{table}

Together, these sources set the frame. Governance law tells deployers to create and preserve records. Evidence law tells reviewers how authenticity and integrity can be established for digital artefacts, without making those properties sufficient for an authorization claim. Agency law clarifies why authority, validity, and accountability have to be separated in AI systems.

Activity-based governance adds reachability, recovery, and cessation without answering the action-level authorization question. Public-sector failures show why reconstruction, a warranted decision basis, routed adverse signals, and effective contestability all matter. Evidentiary assurance joins these materials around a single action-level question: what evidence is needed to evaluate the authorization claim for this action, by this system, under this delegation, at this time?
