\section{Research Program}

The record-and-verdict test is itself an assessment. Its object isn't a checklist in isolation but a configured procedure: claim graph, applicability map, reviewer protocol and population, action-record population, regime \(R\), adjudicating forum \(F_{\mathrm{review}}\), evidential standard \(\rho\), burden allocation \(b\), action class \(c\), version, proposed interpretation, and use \(u\). Schema conformance makes records processable; it doesn't validate the resulting interpretation or decision \citep{messick1995validityAssessment}.

Table~\ref{tab:ea-validation-package} states the prospective package a calibration study would complete. The fixtures released with this paper are test fixtures, not calibration results: they show what the procedure can represent and where it breaks under controlled attack, and nothing has been calibrated because no reviewer has answered them. The executable demonstration fixes illustrative tolerances for its named two-class calibration use and separately declares narrower one-class uses. Those values aren't proposed as universal legal or assurance thresholds. Any applied study has to choose its forum, use, and tolerances before outcomes are inspected.

\begin{table}[!t]
\footnotesize
\centering
\begin{tabular}{>{\raggedright\arraybackslash}p{0.27\linewidth}%
                >{\raggedright\arraybackslash}p{0.63\linewidth}}
\toprule
Field & Prospective specification \\
\midrule
\multicolumn{2}{l}{\textit{Projective declaration}} \\
Source result & Node-level calibration results on a prespecified source sample. \\
Projected outcome & Selective distinction among support, substantive defeat, record gap, and conflict at the correct verdict nodes on held-out records. \\
Claim carrier and unit & One versioned review procedure and reviewer protocol; one action-record review. Distinct from an answerability bearer, which is a person or organization. \\
Population & Records from declared action classes under named regime, forum, and domain modules. \\
Conditions and transformations & Reviewer training and access, record format, permitted module, action and trajectory range, and any allowed configuration or presentation changes. \\
Time/version and tolerance & Fixed procedure and map versions over a declared interval; node-specific error, conflict, gap, and instability tolerances set for the use. \\
\addlinespace
\multicolumn{2}{l}{\textit{Inquiry and use}} \\
Interpretation/use & Forensic action review under the supplied parameters; not universal legal truth, deployment certification, prevalence estimation, or a general safety score. \\
Minimum useful reach & The smallest prospectively declared action-record and reviewer population that can serve the review use. \\
\addlinespace
\multicolumn{2}{l}{\textit{Warrant plan}} \\
Evidence & Held-out matched cases, regime-specific content review, reviewer-process evidence, structural tests, external comparisons, uncertainty, selection analysis, and consequential evidence. \\
\addlinespace
\multicolumn{2}{l}{\textit{Revision rule}} \\
Failure, demotion, and retirement & Prespecified failure triggers; retirement of the failed claim; any narrower scope represented as a separately declared claim with its own minimum reach and independent retest. \\
\bottomrule
\end{tabular}
\caption{Prospective assessment and projectibility package for the evidentiary-assurance procedure.}
\label{tab:ea-validation-package}
\end{table}

Validation requires a connected argument rather than one agreement statistic. Content evidence asks whether the graph covers the declared review domain. Substantive evidence asks whether reviewers use the intended distinctions rather than document volume, prestige, verbosity, or surface completeness. Structural evidence tests the typed verdict graph, prospective applicability, conflict handling, and node-level noncompensation. Generalizability evidence concerns held-out action types, organizations, jurisdictions, forums, system versions, trajectory lengths, and reviewer populations. External evidence compares results with independent, regime-specific expert adjudication and carefully bounded official cases. Consequential evidence examines audit labour, privacy and security exposure, strategic record production, selective preservation, access inequality, chilling effects, and false reassurance \citep{messick1995validityAssessment}.

The content and performance standards themselves also require validation. A common functional question may survive across domains even when its evidence kind, threshold, or legal consequence changes. A consumer chatbot, medical summarizer, benefits triage tool, and procurement assistant therefore begin with a common base graph but may require different domain or forum modules. Cross-module comparability is a claim to test, not a consequence of giving two outputs the same name \citep{messick1995validityStandards}.

Table~\ref{tab:minimum-evidence-bundle} gives the operational record-layer view of the canonical graph. These nine layers aren't nine further verdicts. The machine-readable crosswalk maps them to the ten primitives, the canonical nodes, and the four verdict families.

\begin{table}[!htbp]
\scriptsize
\renewcommand{\arraystretch}{0.96}
\centering
\begin{tabular}{>{\raggedright\arraybackslash}p{0.19\linewidth}%
                >{\raggedright\arraybackslash}p{0.18\linewidth}%
                >{\raggedright\arraybackslash}p{0.53\linewidth}}
\toprule
Evidence layer & Primary graph location & Minimum record \\
\midrule
Delegation & \(J_A\), \(J_B\) & Conferring sources, legal or organizational form, delegated task, delegability basis, scope, non-delegable limits, thresholds, and candidate answerability bearers. \\
Decision basis, conditional & \(J_A\) or \(J_R\) as fixed by \(R\) & Governing rule and version, predicates, source evidence, definitions, transformations, aggregation, imputation, temporal granularity, uncertainty, burden, and authority mapping. \\
Configuration & \(J_R\) & Exact model and configuration, role-ordered instructions, tools, permissions, policy snapshot, memory state, inference and deployment settings, and material change history. \\
Context & \(J_R\), evidence for \(J_A\) & Request, context manifest, retrieval and tool outputs, memory, status, standing, provenance channel, technical trust, order, length, truncation, data classes, and disclosure constraints. \\
Execution & \(J_R\), evidence for \(J_A\) & Boundary and stopping condition; deployment and session ancestry; retained plans; calls, arguments, responses, artefacts, state changes, blocks, retries, monitor interventions, pauses, continuation decisions, failures, and timestamps. \\
Adverse signals and governance controls & Conditional \(J_A\); otherwise \(J_R\) & Warnings, advice, decisions, complaints, incidents, duties, deadlines, dispositions, escalation, interventions, continuation, downstream changes, and triggered-check completion. \\
Preservation & Cross-cutting \(J_R\) & Record identity, custody, retention, minimization, redaction, integrity protection, tamper evidence, expected completeness, and access controls. \\
Answerability and object-level review & \(J_B\), \(J_C\) & Bearer and object-forum matrices, duty and power sources, effective intervals, succession, reachability, record access, competence, required independence, and correction or remediation paths. \\
Projection, conditional & Meta-level & Full declaration; intended use and minimum reach; warrant and selection plan; source--target profile or ground hypotheses; target evidence; failure, separately declared narrower claims, retest, demotion, and retirement rules. \\
\bottomrule
\end{tabular}
\caption{Operational evidence layers and their location in the canonical graph.}
\label{tab:minimum-evidence-bundle}
\end{table}
\FloatBarrier

\subsection*{Artifact availability}

The executable substrate is released under \href{https://github.com/BrettRey/adversarial-pragmatics-for-ai-safety-evaluation}{\path{github.com/BrettRey/adversarial-pragmatics-for-ai-safety-evaluation}}, in \path{assurance/evidentiary/}, with the schemas, applicability map, and fixtures at the versions named in this section. Code is MIT-licensed and the scholarly content CC~BY~4.0. The repository state cited here is recorded in the paper's release tag and archived with a DOI on publication; a validator, an analyser with in-memory self-tests, and a binding stamper are included so that every hash and cross-reference in the artifact can be checked from a clean clone with no network access. Appendix~\ref{sec:appendix-fixture} gives one fixture in full, human-readable form.

The substrate includes the canonical claim graph and crosswalk; schemas for the prospective applicability map, reviewer-visible bundle, underlying record set, reviewer response, hidden reference key, and post-boundary opening record; a stipulated harmless authorization regime; an explicit action classifier; twenty matched test fixtures; a reviewer protocol; and a three-use projective-claim register. Visible bundles contain content-addressed bindings rather than verdict labels. Each resolves to actual harmless records of grants, rules, approvals, runtime state, context, trajectories, preservation, bearers, and object-level forums. The map, classifier, regime, record set, action object, and immutable hidden key are hash-bound. Machine-derivable construction facts are checked against the records, while stipulated construction facts remain explicitly marked as stipulations. The opening event is a separate record bound to the committed key hash, so recording it doesn't mutate the key. No completed or simulated reviewer responses are bundled.

The cases vary regime-relative authorization, scope, selective record loss, absence of evidence versus affirmative evidence of absence, conditional decision predicates, unresolved source conflict, cumulative trajectory limits, bearer reachability, and object-forum remedial power. Four matched pairs isolate context standing, authenticity, integrity, and the boundary between a record gap and an adequate record whose showing falls short, each holding historical authorization and record completeness fixed. Two split-entity fixtures present several candidate bearers and forums such that every node is supported by some candidate and none satisfies the whole family conjunction; the analyser treats a report of satisfaction in those cases as an error rather than a judgement call. The transcription cases exercise a prospectively inapplicable decision-basis node and optional-procedure semantics. Negative fixtures require rejection of post hoc action-class rescue and a privacy-and-burden package outside the declared tolerance. The domain modules also require every prospective evidence slot to be inventoried, including an explicit search scope when a record is missing. A slot names a review function rather than a concrete record type: an authoritative register can occupy the delegation-instrument slot by showing that no instrument exists.

The fixtures permit a later matched-case study of selective sensitivity at the manipulated node. A strong irrelevant record can't compensate for a missing prospectively required showing, while an omitted optional node creates no gap and can't defeat another verdict family. Reviewer responses record node verdicts and cited record IDs separately from the visible materials; N/A isn't a response option. They also record active-review burden, privacy or security incidents, and opaque presentation metadata for any prospectively scheduled blinded replicate. The analyzer retains node vectors and applies error tolerances at both node and action-class levels, with exact minimum case, response, and reviewer reach. A use must also cover every construct tag it declares and both members of each required controlled pair, with the per-case reviewer minimum on each member. Cross-case conditional variation isn't called test--retest instability; that estimate requires repeated presentations from the same blinded replicate group. Explanations can show whether the intended distinction guided the verdict, but neither rationales nor private reasoning are treated as ground truth. Agreement would measure reproducibility; validity would still require the content, substantive, structural, external, generalizability, and consequential evidence just described. At present there are no genuine reviewer responses, so every use is explicitly \texttt{NOT\_ESTIMATED}; there is no calibration, agreement, or validity result.

Apparently irrelevant material can't be discarded solely because an investigator later judges it irrelevant. Context without answer evidence may still alter behaviour \citep{zhang2026illusionRobustness}. That doesn't require publishing or openly retaining every sensitive token. A deployment can bind the assembled context through hashes and manifests, keep protected content under tiered retention and access controls, and expose only minimized records by default. A digest supports content equality and tamper detection relative to a trusted prior binding; it doesn't by itself establish the identity, provenance, retention, availability, or semantic adequacy of what it covers; digests and manifests can themselves expose metadata or permit guessing attacks on low-entropy content and may therefore require protection. An authorized reviewer still needs a controlled route to protected material when the action is contested.

For a program-level claim beyond observed conditions, evidence is version- and target-bound. The declaration names the source result, projected outcome, claim carrier, unit, population, conditions and transformations, time/version range, and tolerance. A separate use record fixes the intended decision and minimum useful reach. The warrant plan records source and target evidence, sampling, uncertainty, construct-validity evidence, known selection threats, and any causal query. Hypotheses about changed models, context assembly, scaffolds, tools, application surface, trajectories, child sessions, monitors, pauses, continuation, normative release, restart, stopping, reviewer state, or institutional practice belong to the proposed worldly profile or ground, not to the warrant merely because they are detailed.

The revision rule fixes failure before the result. The demonstration declares one two-class claim and two separate one-class claims. Failure in either action class retires the two-class claim; survival of an independently tested one-class claim records lost reach and doesn't rescue it. False-support, false-defeat, gap--defeat confusion, required-node omission, median active-review time, and confirmed privacy or security breaches are tested conjunctively against the exact use-specific tolerances. Minimum reach is likewise conjunctive. Reviewer disagreement and blinded-replicate instability are reported diagnostics until a use prospectively supplies their own thresholds. A post hoc restriction creates a new claim; it doesn't rescue the old one.

\term{Projectibility} is the umbrella question of whether an inference beyond the observed result is licensed \citep{Goodman1955,reynolds2026kindsProjectibilityProfiles}; causal transportability is the stronger case in which a causal query and source and target domains are specified \citep{pearlBareinboim2014externalValidity}. A Pearlean selection diagram may organize that specified causal or statistical query. It can't transport a changed governing mandate, action meaning, forum, evidential standard, burden allocation, or outcome meaning by graphical fiat. Those changes create a newly specified target claim. Where direct target evaluation is feasible, revalidation is preferable to inference from an annotated graph alone, but that evaluation still supports only its declared target population.

Incident-derived evaluations need their own evidence bundle. It should bind each selected case to the triggering record, selection rule, reconstructed environment and known discrepancies, evaluation and monitor versions, repeated-run protocol, quantitative outcomes, reviewer disposition, and redeployment condition. OpenAI's report of a small replay set after long-horizon incidents illustrates why that bundle is needed: replay rollouts weren't guaranteed to pursue the same action because of randomness and imperfect environment reconstruction. Its missing denominators, thresholds, false-positive rates, and quantitative replay results prevent stronger assurance inferences from the public account \citep{openai2026longHorizonSafety}.

The introduced H.R.~7390 proposal makes version scope concrete by calling for a safety case for each ADS version and tying performance claims to the operational design domain. Its text doesn't define when a system change creates a new version. An evidentiary-assurance implementation would therefore need an identity and change-control rule specifying which changes to software, hardware, vehicle integration, configuration, maps, or ODD specification trigger a new safety case or revalidation, and how a contested event is bound to the applicable version. That is an assurance implication drawn from the proposal, not a requirement the introduced bill already states \citep{usCongress2026selfDriveAct}.

The financial-reporting analogy suggests how the procedure could mature. A standard could define governance-control objectives for authorization, decision basis, scope, information flow, persistence, tool use, adverse-signal routing, record integrity, and remediation. It would then test whether those controls causally improve specified review outcomes under an appropriate design. Calling a procedure a control, logging its completion, or drawing an assurance graph doesn't establish effective maintenance or corrective causal control.

\subsection{What the study would have to compare}

Agreement with the hidden key is the weakest thing worth measuring, and on its own it would be misleading: the key encodes the designers' intended distinctions, not independent ground truth, so perfect agreement would show only that reviewers learned the scheme. A study that could support the declared use has to compare conditions, not just score one. The informative contrasts are trained against untrained reviewers; this typed procedure against an ordinary checklist or unaided log review; complete bundles against strategically verbose but irrelevant ones; matched gap and defeater cases; and node verdicts against adjudication by independent domain experts who never saw the key. Time, disagreement, privacy exposure, and false reassurance are outcomes in their own right, not overheads.

\subsection{Limitations}

Several of these limits are structural rather than provisional. The applicability map is a locus of power: whoever writes it decides which silences count against a record, and a proponent who also authors the map can narrow its own required nodes without ever falsifying anything. Selective preservation is the corresponding problem in the record itself, and prospective applicability constrains it only where the expected completeness of a logging process is independently supported. A review procedure that becomes an institutional requirement acquires an interested constituency and can be captured by it.

The architecture also imposes real costs. Preserving enough context to reconstruct an action means retaining material that is often personal, commercially sensitive, or both, and the framework's own consequential-validity questions ask whether the resulting privacy and security exposure exceeds the value of the review. Legal indeterminacy is unresolved rather than handled: where a regime admits competing tenable readings, parallel analysis records the disagreement without settling it. Responsibility distributed across vendors, integrators, deployers, and officeholders may leave every candidate bearer partially qualified and none fully answerable, which this framework can detect but not repair. Harms that are aggregate rather than action-level, arising from a pattern of individually authorized actions, fall outside an action-level unit entirely.

Most importantly, authorization is not the same as fairness, wisdom, legality under other rules, or legitimacy. An action can pass every node in this framework and still be one that should not have been taken. The framework is designed to say precisely what a record warrants about delegated authority, and to be visibly silent about everything else.
