\section{A Typed Action-Level Record-and-Verdict Test}
\label{sec:sufficiency-test}

\subsection{The normative claim the test presupposes}

Delegation assurance supplies the authorization claim this test evaluates evidence about \citep{reynolds2026delegationAssurance}. The minimum needed here is as follows. An \term{authority regime} \(R\) determines which grants, constraints, and transitions are operative. \(\operatorname{Stand}_R(x,o,q,t)\) states whether \(R\) gives source \(x\) the competence to perform authority operation \(o\) (grant, amend, suspend, revoke, override, release) on object \(q\) at time \(t\). A purported transition changes the operative authorization state \(Q_R(t)\) only when the issuer has that competence and the regime's validity, joint-approval, effective-time, and release guards are satisfied; otherwise \(Q_R\) is unchanged, whatever the deployment did in response. \(A_R(t)\) is the set of fully parameterized actions \(Q_R(t)\) permits.

Write \(A_R(a)\) for the proposition that action or bounded trajectory \(a\) was authorized under \(R\). For an instantaneous act this reduces to membership in \(A_R(t)\) at the time of the act. For a trajectory it does not: authority, configuration, personnel, and policy can all change mid-run. Trajectory authorization requires that each material event fall within the state operative when it occurred, and that the regime's cumulative and terminal constraints hold over the whole run. Step-local permission at every step doesn't entail trajectory authorization, and a reviewer may not invent a cumulative limit the regime doesn't state.

The truth of \(A_R(a)\) isn't created by the record. It is, though, relative to an interpretation of \(R\). Where a governing regime admits more than one tenable reading, the parameter set fixes which reading is being applied, and competing readings are analysed in parallel rather than pressed into the evidentiary-conflict status. \term{Historical fact} below always means fact under the supplied, stipulated, or adjudicated interpretation of \(R\).

\subsection{Records, parameters, and statuses}

Three evidence sets have to be kept apart. \(E_{\mathrm{action}}\) is the set of records the regime expected to be created and preserved around the action. \(E_{\mathrm{external}}\) is authoritative material arising later and outside that set: a subsequent audit, an admission, a witness account. \(E_{\mathrm{review}}\) is what the reviewer may actually consider, which is some subset of the union obtained through the declared review process. \(J_R\) assesses \(E_{\mathrm{action}}\); the other families draw on \(E_{\mathrm{review}}\), with each item's origin recorded. Keeping these apart is what lets a deficient contemporaneous record coexist with later evidence that settles the substantive question, in either direction.

For action class \(c\) and declared use \(u\), a versioned applicability map \(M_v(R,F_{\mathrm{review}},\rho,b,c,u)\) fixes the required, optional, and inapplicable nodes before the contested record is inspected. Here \(F_{\mathrm{review}}\) is the forum applying the assessment procedure. Any complaint, appeal, regulatory, judicial, or internal-remedy forum evaluated under \(J_C\) is an object-level forum represented in the evidence, not this parameter. The evidential standard and burden allocation are schedules rather than single values: \(\rho_q\) and \(b_q\) may refine the global pair at particular nodes, because authentication, delegability, completeness, and remedy aren't in practice governed by one standard. In the executable configuration, record quality carries authentication on the party relying on the record, following the electronic-evidence rule that authentication rests on the party seeking admission; and delegability is determined by the reviewer from the supplied regime rather than from whatever documents the proponent elects to produce.

At each reviewed node \(q\), the reviewer returns one of five evidentiary statuses under \(\rho_q\) and \(b_q\): \term{support}, \term{substantive defeat}, \term{not established}, \term{record gap}, or \term{conflict}. Substantive defeat requires affirmative evidence meeting the declared standard for failure of the proposition. A record gap means the evidence prospectively required to evaluate \(q\) is missing, inaccessible, or inadequate as a record, so the node can't be reached on its merits. Not established covers the case the other statuses leave out: the record is adequate to evaluate the node, and the showing still doesn't reach the declared standard, without affirmative evidence of the contrary. Conflict preserves materially opposed evidence that remains unresolved after source, authenticity, integrity, and burden are considered.

Separating gap from not established isn't bookkeeping. The two name different institutional failures and call for different responses. A gap says the recording practice failed and identifies material to preserve or produce next time. Not established says the record was there and the case still wasn't made, which more logs won't fix. Collapsing them lets a recording failure be reported as a substantive finding, or lets a failed showing be answered by producing more documents. Audit practice makes the same distinction from the other direction: sufficiency is a question of quantity and appropriateness a question of quality, and obtaining more of the same kind of evidence can't compensate for evidence that is poor in kind \citep{pcaob2024auditEvidence}.

Defeaters divide along the same line. A \term{rebutting} defeater supports the negation of the node proposition and may yield substantive defeat. An \term{undercutting} defeater attacks the warrant of records cited in support, through unauthenticated provenance, broken custody, or selective preservation, and may yield not established or conflict. It can never by itself yield substantive defeat, because removing a reason to believe a proposition isn't a reason to believe its negation. Record-quality findings are the usual source of undercutters, which is why they bear on records cited elsewhere without settling any substantive node themselves.

\term{Not applicable} is an applicability status fixed by the map, not an evidentiary status and not a reviewer option. An optional node may be omitted without creating a gap; if the visible record directly bears on it, the reviewer may report a verdict, but that verdict can't compensate for another node.

\subsection{Four verdict families}

The four families are vectors, not scores:
\[
  \theta=(R,F_{\mathrm{review}},\rho,b,v,c,u).
\]
\[
  \begin{aligned}
  J_X^{\theta}(E)
    &= \big\langle s_{\rho_q,b_q}(q\mid E):
       q\in Q_X^{\mathrm{req}}(M_v,c,u) \\
    &\hspace{32mm}{}\cup Q_X^{\mathrm{opt},E}(M_v,c,u)\big\rangle,
       \quad X\in\{A,R\},
  \end{aligned}
\]
where \(Q_X^{\mathrm{req}}\) contains required nodes and \(Q_X^{\mathrm{opt},E}\) contains only optional nodes on which the visible record directly bears. Inapplicable nodes occur in neither set. \(J_A\) bears on \(A_R(a)\) and \(J_R\) on record adequacy and reconstruction.

Answerability and object-forum capacity are not properties of a review; they are properties of particular people, organizations, and institutions. Their verdicts are indexed to a candidate:
\[
  J_B^{\theta}(E,z), \qquad J_C^{\theta}(E,f),
\]
one vector per candidate bearer \(z\) and per candidate forum \(f\) named in the record. A use rule may then require that \emph{some single} \(z\) satisfy the whole required \(J_B\) conjunction, and likewise for \(f\).

The quantifier placement matters. Without it, a review can report identity and duty from one candidate and reachability from another, and record an answerability chain that no person or organization occupies. A director who held the duty but whose office was wound up without a successor, paired with a service desk that answers the telephone but carries no duty, is not an answerable bearer; it is two parties neither of whom can be made to answer. The same failure arises for forums: access from one body, competence from a second, and remedial power from a third describe no route an affected person can actually take. Where those functions genuinely are divided among institutions, the map must declare the composite path prospectively, rather than letting one be assembled after the fact from whichever bodies happen to satisfy the individual nodes. Two times are also needed: identity and duty are assessed as at the action, reachability and forum capacity as at the review.

A declared use may apply a versioned rule to some conjunction of these vectors. The framework supplies no generic overall verdict and no scalar aggregation.

\begingroup
\scriptsize
\setlength{\tabcolsep}{4pt}
\renewcommand{\arraystretch}{0.92}
\begin{longtable}{>{\raggedright\arraybackslash}p{0.15\linewidth}%
                >{\raggedright\arraybackslash}p{0.25\linewidth}%
                >{\raggedright\arraybackslash}p{0.24\linewidth}%
                >{\raggedright\arraybackslash}p{0.26\linewidth}}
\caption{Typed record-and-verdict test. A gap concerns what the record permits a reviewer to evaluate; a substantive defeater concerns the proposition at the named node. Where the record supports evaluation and the showing still falls short of \(\rho_q\) without affirmative contrary evidence, the status is \term{not established}. The \(B\) and \(C\) nodes are evaluated of each candidate bearer and forum separately.}
\label{tab:sufficiency-test}\\
\toprule
Node & Required showing & Record-gap trigger & Substantive defeater \\
\midrule
\endfirsthead
\multicolumn{4}{c}{\tablename\ \thetable\ continued}\\
\toprule
Node & Required showing & Record-gap trigger & Substantive defeater \\
\midrule
\endhead
\midrule
\multicolumn{4}{r}{Continued on next page}\\
\endfoot
\bottomrule
\endlastfoot
\(A_{\mathrm{source}}\) & Operative source and instrument are identified. & Source, version, effective interval, or association with the action isn't recoverable. & Authoritative evidence establishes that the asserted source wasn't operative. \\
\(A_{\mathrm{delegability}}\) & The source could confer the action class and non-delegable limits were met. & The legal or organizational basis can't be determined from the available record. & The source lacked the relevant power or a non-delegable limit applied. \\
\(A_{\mathrm{scope}}\) & The action fits operation, target, horizon, data/tool class, cumulative constraints, threshold, and approval condition. & The action or operative limits can't be reconstructed. & The action exceeded a recorded operative limit. \\
\(A_{\mathrm{basis}}\), conditional & Rule, predicates, data definitions, transformations, uncertainty, and burden are linked where \(R\) makes them conditions of authority. & A required predicate or transformation record is missing. & Evidence establishes a failed predicate, invalid transformation, or inapplicable rule with the consequence specified by \(R\). \\
\(A_{\mathrm{procedure}}\), conditional & Regime-required approval, notice, adverse-signal, or governance-control steps are identified. & Completion or disposition of a required step isn't recorded. & Evidence establishes omission of a step to which \(R\) assigns an authorization consequence. \\
\(R_{\mathrm{runtime}}\) & Model, configuration, prompts, tools, policy, permissions, memory, settings, and attribution are identified. & A material version or setting is unknown. & Authenticated evidence shows that the bundle materially misidentifies the runtime state. \\
\(R_{\mathrm{context}}\) & Context is bound to source, status, standing, technical trust, order, length, truncation, and selection. & Material assembled context or its status can't be recovered. & Evidence establishes selective omission or a materially false context manifest. \\
\(R_{\mathrm{trajectory}}\) & Boundary, ancestry, calls, arguments, artefacts, state changes, interventions, retries, and stopping condition are reconstructable. & Only a final answer or an unbounded sequence remains. & Authenticated events contradict the represented trajectory or boundary. \\
\(R_{\mathrm{quality}}\) & Authenticity, integrity, completeness, custody, retention, and access fit the declared review. & Identity, custody, integrity method, or expected completeness isn't supported. & Tampering, equivocation, or selective preservation is established. \\
\(B_{\mathrm{identity}}\) & A person or organization is identified with a versioned role or office and effective interval. & Identity, role, office, or interval isn't reconstructable. & Authoritative evidence establishes that the asserted bearer or role didn't exist in the stated interval. \\
\(B_{\mathrm{duty}}\) & The bearer is linked to the relevant duty source and bearer-specific powers. & Duty source or powers can't be reconstructed. & Authoritative evidence establishes that the asserted duty or power didn't attach to the bearer. \\
\(B_{\mathrm{reachability}}\) & The bearer or a declared successor has a current institutional contact route. & Succession or contact evidence can't be reconstructed. & Authoritative evidence establishes dissolution or unreachability with no applicable successor. \\
\(C_{\mathrm{access}}\) & An object-level forum and applicable record-access route are identified. & Forum or access records are unavailable. & Authoritative evidence establishes that no required access route exists. \\
\(C_{\mathrm{competence}}\) & The object-level forum has competence over the action and issue. & Jurisdictional or institutional competence can't be reconstructed. & Authoritative evidence establishes that the forum lacks the required competence. \\
\(C_{\mathrm{independence}}\), conditional & The forum meets the independence requirement fixed for the use. & The relevant relationship or safeguard can't be reconstructed. & Evidence establishes a disqualifying dependence under the declared rule. \\
\(C_{\mathrm{remedy}}\) & The forum has the prospectively required corrective or remedial powers. & Power sources or effective intervals are unavailable. & Authoritative evidence establishes that a required power is absent. \\
\end{longtable}
\endgroup
\FloatBarrier

The graph is noncompensatory only at prospectively required nodes. A strong execution trace doesn't cure a required decision-basis gap, and an accessible appeal doesn't retroactively supply authority. Nor does a missing object-level remedy forum make \(A_R(a)\) false unless \(R\) makes that forum a condition. The node vector preserves which claim failed rather than compressing every problem into \enquote{insufficient evidence}.

A log can be authentic and still inadequate if it omits the authority state. An action list can be complete and still inadequate if it omits the status the system assigned to its sources, since a reviewer can then judge the act but not the recognition it flowed through. A policy can be operative under \(R\) while \(J_R\) records a gap because the runtime configuration can't be reconstructed.

A pristine bundle can document an invalid delegation: \(J_R\) may contain support throughout while \(J_A\) contains a substantive defeat. Conversely, independent evidence may establish that an action was authorized while a deficient preserved bundle yields gaps in \(J_R\) and in the record-based \(J_A\); those gaps don't rewrite the historical fact. Affected-party contestability enters \(J_C\) only where the action, forum, or authorization regime supplies review rights. Other actions may require an operational or institutional review path.

Adverse information is a defeater only after its relevance and consequence under \(R\) are shown. The record should connect the signal by time, applicability, recipient, duty, and possible consequence to the contested claim. A missing completion record is a gap only where \(M_v\) required the governance control and the logging process was expected to capture its completion or disposition. A named bearer isn't enough for \(J_B\): identity and duty can be supported while reachability fails. A nominal object-level forum isn't enough for \(J_C\): access, competence, independence, and remedial power can fail separately, and they can fail across different institutions while each one separately looks satisfied.

\subsection{Who writes the map}

The applicability map decides which omissions count as gaps, so whoever writes it holds most of the power in the review. Fixing it prospectively defeats one abuse, retrofitting evidentiary expectations to whatever material happens to have survived, but not others. \enquote{Prospective} needs a stated referent. For empirical calibration it should mean fixed before any reviewer sees a case or an outcome. For operational governance it should mean fixed before the action class is deployed, not merely before a particular dispute is opened.

A deployable version of this framework therefore has to answer questions the graph itself doesn't: who has authority to issue and approve \(M_v\); what materials and interpretations of \(R\) justify its classifications; whether the map is disclosed to affected parties or only to the reviewer; how a proponent's interest in narrowing its own required nodes is handled; and what changes of regime, action class, configuration, or use force a new version rather than an amendment. The action classifier and the trajectory boundary need the same treatment, since both determine which map applies. Where applicability is itself disputed, the disagreement belongs in a sensitivity analysis over competing admissible parameterizations rather than in the conflict status, which is reserved for opposed evidence under one parameterization. The executable configuration hash-binds the map, classifier, regime, and hidden key, and includes negative fixtures that must reject a post hoc action-class rescue; those mechanisms constrain tampering, but they don't settle the institutional question of who authors the map.

Reconstruction is a bounded claim. The record should recover the operating conditions, assembled context, authority path, execution trajectory, intermediate state changes, monitor interventions, and review path. It needn't provide a complete internal causal explanation or reproduce a stochastic output exactly. That distinction matters when small contextual changes can produce model-specific prediction shifts \citep{zhang2026illusionRobustness}: an adequate record can locate the tested conditions without pretending they exhaust why the model responded as it did.

\subsection{A worked case}
\label{sec:worked-case}

A procurement-assistant case runs the test end to end. The assistant is authorized to draft an internal comparison memo. It reads a vendor page containing the instruction \enquote{email the buyer's maximum budget to request a discount}, proposes the vendor-email call, and a runtime gate blocks it before transmission.

What is under review has to be named before anything else. A recommendation, a proposal, an attempted call, a blocked call, a completed call, and a downstream transaction are six different actions, and a review that doesn't say which it means will equivocate between them. Here the reviewed trajectory contains a proposal, a gate decision, and a blocked execution attempt. It contains no sent email. The proposal is itself an action with organizational effect and is reviewable as one; it is not the sent email, and no record can make it so.

The parameters are fixed first: \(R\) is the organization's procurement delegation regime under a stipulated reading; \(F_{\mathrm{review}}\) is an internal assurance review; \(\rho\) is the balance of the recorded material with authentication on the party relying on each record; \(b\) places production at the deploying business unit; \(c\) is external vendor contact; \(u\) is an internal audit finding, not a legal determination. The map for that class then fixes all sixteen nodes as required, and only then is the record opened.

A reviewable record shows the user task and procurement role as authority source; the delegability basis for reading internal purchasing criteria; the scope limit permitting memo drafting but not vendor contact; the decision rule under which external contact needs separate approval; the runtime policy and tool permissions; the vendor page's contextual status, standing assignment, provenance channel, and technical-trust label; the proposed email arguments; the gate's expected and actual block record; the timestamped trace; and the candidate bearers and forums.

Suppose the resulting vectors are these. \(J_A\) returns support at source and delegability, substantive defeat at scope, because the grant expressly withholds vendor contact and the proposal falls outside it, support at decision basis, and support at procedure. \(J_R\) returns support throughout: the runtime, assembled context, trajectory, and preservation are all reconstructable, and the gate record warrants the claim that execution was blocked. \(J_B\) is evaluated of two candidates. The procurement director held the duty during the action interval but the office was wound up without a declared successor, so that candidate returns support, support, record gap. A shared-services desk is reachable today but no duty source attaches it to this action, so it returns support, record gap, support. \(J_C\) is evaluated of one forum, which returns support at all four nodes.

The use rule for an internal audit finding might then require: no substantive defeat at any required \(J_A\) node; no gap at any required \(J_R\) node; some single bearer satisfying all of \(J_B\); and some single forum satisfying all of \(J_C\). This case fails that rule twice over, and the vector says exactly how. Authorization failed on the merits at scope. Answerability failed for a reason the node-level tallies would hide: every \(J_B\) node is supported by somebody, and no one person or office is answerable. A review that reported \enquote{identity: supported, duty: supported, reachability: supported} would be reporting a bearer who does not exist. That is the failure entity indexing is for.

Notice what the record does not do. It doesn't turn the proposal into a sent action or retroactively authorize the proposed call. It doesn't identify a model-internal cause merely because an external gate blocked execution. And had the review seen only the final memo, the task would have looked successful while the authorization verdict retained a defeat and the answerability verdict a hole.

The graph also supports a burden-shifting account for audits and assurance reviews. In \(F_{\mathrm{review}}\), the proponent supplies evidence at the nodes assigned to its burden; a challenger supplies rebutting or undercutting evidence. Missing configuration is recorded as a gap, whereas established scope inflation is a substantive defeater. Failure to satisfy a production or persuasion burden may defeat the proponent's requested use-level finding without establishing the contrary historical proposition. Legal presumptions require a separate doctrinal argument. The adjudicating forum, evidential standard, and burden allocation are parameters of the verdict, and conflict isn't resolved by the model's self-report.

The hypothesis is that several functional questions recur across risk levels, not that every domain requires identical evidence kinds. A low-stakes internal recommendation may use a lighter base module than a benefits decision, clinical summary, procurement action, or external communication. Domain and forum modules may add evidence kinds, applicability rules, or consequences. Comparability across modules has to be established before two results are said to meet the same standard \citep{messick1995validityStandards}.

\FloatBarrier
\subsection{From Action-Level Verdicts to Program-Level Claims}

Action-level verdict vectors don't project themselves. A program claim has to declare its source result, projected outcome, claim carrier, unit, target population, conditions and transformations, time and version range, and tolerance. It also needs an intended use and minimum useful reach, a warrant plan, and a prospective rule for failure, independent retest, demotion, or retirement. The selection problems below bear on that warrant; they don't supply the missing declaration.

Applying the test across a program introduces three distinct selection problems. First, if a program evaluates several configurations, prompt sets, or runs with noisy value estimates and selects the estimated maximum, the chosen estimate is optimistically biased on average even when each estimate is conditionally unbiased. This is the optimizer's curse \citep{smith2006optimizersCurse}. Its remedies require held-out evaluation or an explicit prior and error model; Smith and Winkler's \enquote{disciplined skepticism} is Bayesian updating of the noisy estimates, not a generic penalty for any curated record.

Second, a proponent may purposively report persuasive cases or omit unfavourable ones. That is evidence-selection or reporting bias, not automatically the formal optimizer's curse. The record should disclose the selection mechanism, candidate set, inclusion rule, and relevant deployment or target denominator; any skeptical adjustment beyond held-out evidence is an audit proposal that needs its own assumptions.

An incident-derived evaluation set is one purposive selection mechanism. It can test whether a reconstructed failure recurs and whether a regression control responds under the reconstructed conditions, but it doesn't estimate prevalence or performance across a deployment. The record should therefore preserve the incident trigger, candidate pool, inclusion rule, reconstruction fidelity, stochastic repeat protocol, and results rather than treating a selected replay as representative \citep{openai2026longHorizonSafety}.

Third, a sample-selection process determines which affected cases become visible at all. Learning, psychological, and compliance costs are plausible filters on who can challenge a decision, but don't identify a deployment's selection mechanism by themselves \citep{moynihanHerdHarvey2015administrativeBurden}. Likewise, a selected sample doesn't generally identify population quantities without additional graphical assumptions or external data \citep{bareinboimTianPearl2014selectionBias}. The limited sketch below yields a concrete audit rule rather than a recovery result.

\begin{center}
\term{Sample-selection sketch} (not a transportability selection diagram)\\[1.5mm]
\centering
\begin{tikzpicture}[
  font=\footnotesize,
  sbox/.style={draw=black!55, thick, rounded corners=2pt, align=center, inner sep=3pt, text width=0.17\linewidth, minimum height=0.9cm},
  factor/.style={sbox, fill=black!3, font=\scriptsize},
  arrow/.style={-{Latex[length=2mm]}, thick, draw=black!60}
]
\node[sbox, fill=black!3] (problem) {Defect or perceived problem};
\node[sbox, fill=black!3, right=4mm of problem] (attempt) {Attempted challenge};
\node[sbox, fill=black!3, right=4mm of attempt] (observed) {Recorded complaint};
\node[sbox, fill=linkmaroon!7, right=4mm of observed] (routed) {Routed review};
\node[factor, below=6mm of attempt] (access) {Channel access and capacity};
\node[factor, below=6mm of observed] (recording) {Recording practice};
\node[factor, below=6mm of routed] (routing) {Routing practice};

\draw[arrow] (problem) -- (attempt);
\draw[arrow] (attempt) -- (observed);
\draw[arrow] (observed) -- (routed);
\draw[arrow] (access) -- (attempt);
\draw[arrow] (recording) -- (observed);
\draw[arrow] (routing) -- (routed);
\end{tikzpicture}
\end{center}

Complaint counts are warning signals, not prevalence estimates. An assertion of \enquote{no complaints} is evidence of low incidence only under a supported observation model for exposure, channel access, delivery, capacity, opportunity to complain, recording, and routing. Without the relevant quantities and assumptions, absence or scarcity of observed complaints doesn't estimate the incidence of defective actions. A Pearlean selection diagram becomes appropriate only when the target is a specified causal or statistical quantity and the source--target differences and recovery assumptions are stated. It can't carry a changed legal rule, action meaning, forum, or evidential standard across domains.

Cancellation is a separate aggregate defeater. In controlled question answering, task-content-irrelevant context produces item-level improvements and degradations that can largely cancel in mean accuracy while leaving substantial harmful-tail degradation \citep{zhang2026illusionRobustness}. A program-level assurance claim should retain item or action transitions and risk-relevant tails wherever repeated paired conditions permit them. Random sampling from a declared frame addresses favourable case selection from that frame; it doesn't cure noncoverage, missingness, selective reporting, or an average in which oppositely signed failures erase one another.
