\section{Conclusion}

The literature disperses evidence across neighbouring fields. Explainability helps with reasons, provenance helps with lineage, audit logging helps with reconstruction, governance law creates documentation duties, electronic-evidence law supplies concepts for authenticity and integrity, and the assurance-case tradition supplies structure for linking claims to evidence while explicitly declining to say when the contents are good enough. None of these materials alone answers the delegated-authority question, and none makes clean execution evidence a substitute for a warranted decision basis, routed adverse signals, prospectively required governance controls, or meaningful review.

Evidentiary assurance integrates those materials around a bounded action-level review. Delegation assurance supplies the normative claim \(A_R(a)\) and transition semantics. Evidentiary assurance supplies four evidence-relative verdict vectors, parameterized by regime, adjudicating review forum, evidential standard, burden allocation, applicability-map version, action class, and declared use: \(J_A\) for what the record warrants about authorization conditions, \(J_R\) for record adequacy, \(J_B(z)\) for the identity, duty, and reachability of a named bearer, and \(J_C(f)\) for the access, competence, independence, and remedial power of a named forum. Each preserves support, substantive defeat, a showing that falls short on an adequate record, a record gap, and unresolved conflict at the node where it arises.

Historical authorization stays distinct from what a record warrants about it, so poor recording doesn't make an authorized action unauthorized and pristine recording doesn't cure invalid authority. A record that can't be evaluated stays distinct from one that can be and doesn't persuade, because the first calls for better preservation and the second won't be helped by it. Answerability holds of a person or organization rather than of a review, so a chain assembled from several partially qualified candidates identifies nobody who can be made to answer.

The framework supplies no universal assurance score. The executable fixtures show that the distinctions can be encoded and exposed to controlled failure; with no reviewer responses yet, they don't show that reviewers can apply them reproducibly or validly. The proposed standard remains a research hypothesis until its interpretation and use are supported by content, substantive, structural, generalizability, external, and consequential validity evidence, followed by target-bound projection and prospective withdrawal when its fixed minimum reach is lost. Those questions remain subject to forum- and jurisdiction-specific determination. The account is narrower than a finding of full legality, fairness, wisdom, legal liability, moral responsibility, or substantive correctness.
