\section{What Existing Work on Control and Security Leaves Open}

Two literatures bear on delegated machine action. One supplies threat models, benchmarks, control mechanisms, audit proposals, and assurance methods. The other specifies when human control over an autonomous system counts as meaningful, when oversight is effective, and what accountability requires. Neither settles whether a concrete operation was authorized, and the gap between them is where this paper sits.

\subsection{Agent Security and Prompt Injection}

Agent-security guidance has moved beyond ordinary chatbot safety. OWASP's agentic-AI guidance frames autonomous systems through threat models and mitigations \parencite{owasp2025agentic}. Work by NIST and CAISI asks how to measure, constrain, and monitor agent access in deployment environments \parencite{nistCaisi2026agents}. This literature identifies the relevant attack surface: planning, tool use, memory, external content, credentials, and interaction with ordinary software.

Prompt-injection research gives the sharpest technical motivation. Indirect prompt injection shows that LLM-integrated applications can confuse untrusted data with instructions when external content is retrieved or processed inside a larger application \parencite{greshake2023not}. AgentDojo turns that vulnerability into a dynamic evaluation environment for tool-using agents, with realistic tasks, attacks, defences, and security test cases \parencite{debenedetti2024agentdojo}.

These works establish the security problem. They don't by themselves supply typed semantics for whether a concrete operation or trajectory was authorized, a representation that preserves the governing regime's dependency structure, or diagnostics and comparative tests for the path from governing language to execution.

\subsection{Instruction Hierarchy and Language-Mediated Authority}

Instruction-data-separation work treats embedded directive-form content as a status-classification problem: a model should act on its governing instruction while processing, rather than treating as controlling, content embedded in data that purports to instruct \parencite{zverev2025separate}. Instruction-hierarchy work treats prompt injection as a priority problem: systems have to follow privileged instructions over lower-priority or untrusted strings, and training can improve selective ignoring of lower-priority content \parencite{wallace2024instruction}.

Delegation assurance adds authorization standing, or transition competence: message channel, storage location, and technical trust bear on whether a source can confer or alter authority only insofar as the governing regime recognizes them. These questions matter because retrieved documents, memories, policy examples, and tool outputs may contain directive-form strings without conferring authority.

Contextual status identifies how a string occurrence functions; standing identifies whether its source can confer or alter authority; priority resolves conflicts among eligible directives; and licensing connects those determinations to action authorization. Delegation assurance needs separate records of the system's status classification, standing assignment, applied conflict rule, and resulting authorization verdict.

\subsection{Human Control, Oversight, and Accountability}

The closest normative literature asks when human control over an autonomous system is meaningful. \Textcite{santoniDeSio2018meaningfulHumanControl} give two conditions. A system is under meaningful human control when its decision-making tracks the moral reasons relevant in the circumstances, and when its actions and states trace to the proper understanding of at least one human in the design or use context. They note explicitly that the tracked reasons needn't be the operator's: designers, programmers, legislators, or policy-makers may supply them.

That last point is where an authority regime becomes necessary. Tracking some human's reasons along the chain leaves open which human's reasons the regime makes operative, for which operation, at which time. A system can track a designer's reasons faithfully while executing an operation the designer had no competence to license.

\Textcite{sterz2024effectiveOversight} specify effectiveness for the overseeing person: causal power to establish a sufficient causal connection to the relevant aspect of the world, epistemic access to the decision situation, self-control to decide and follow through, and intentions fitting for the role. \Textcite{novelli2024accountability} decompose accountability into context, range, agent, forum, standard, process, and implications, and separate its goals of compliance, report, oversight, and enforcement.

These are conditions on persons and on relations between persons. None is a condition on authority. An overseer may have causal power, epistemic access, self-control, and fitting intentions while lacking standing to amend the mandate; a source may have standing to amend it while satisfying none of the four. Novelli et al.'s agent-and-forum structure identifies who answers for an action after the fact, which is distinct from who could authorize it beforehand. Section~\ref{sec:delegation-beyond-responsiveness} keeps answerability and transition competence apart for that reason.

\Textcite{nguyen2026controllability} survey controllability mechanisms for agentic AI and draw the distinction this paper needs, between causal controllability, the technical capacity to intervene, and normative controllability, the conditions under which intervention constitutes genuine oversight. Delegation assurance accepts that distinction and adds an operational requirement: the two have to be separately recorded in the trace, because a deployment can be causally interruptible at every step while no record shows which source was competent to license the operation. A conceptual distinction that leaves no evidential trace can't be audited.

\subsection{Delegation Protocols and Access Control}

The closest protocol work already uses delegation language. \Textcite{south2025authenticated} propose authenticated, authorized, and auditable delegation for AI agents by extending OAuth/OpenID-style infrastructure with agent credentials, scoped permissions, accountability chains, and natural-language permission translation. \Textcite{kuehlewind2026audit} approach the same problem from the audit side in an IETF Internet-Draft, a work-in-progress rather than a standard, linking user intent, delegation, authorization, execution, and trace reconstruction through audit records and propagated audit context. Newer agent-security mechanisms sharpen the per-call question: ScopeGate distinguishes tool exposure from per-call authorization \parencite{mellafezuvic2026scopegate}, and AttriGuard uses counterfactual causal attribution to separate tool calls supported by a user request from calls driven by untrusted observations \parencite{he2026attriguard}. The framework below uses both as comparison conditions.

This work supplies core infrastructure for delegation assurance. Scoped credentials, delegated principal identity, authorization-transition records, and audit context can make recorded authorization states more explicit than ordinary prompts or application logs. The remaining gap lies in joining delegation infrastructure to language-status classification, refusal calibration, evaluator-label use, authorization-claim fields, and comparative evaluation.

A protocol can propagate a recorded scope without establishing whether the classification of governing language fit its contextual status, whether the action path separated read access from disclosure authority, whether a refusal was compatible with \(R\), whether completion was required, or whether a downstream decision rule used a validated evaluator label within its declared scope.

\subsection{Assurance Cases, Governance, and Provenance}

Assurance-case methods provide the right argumentative shape. The claim/argument/evidence structure in argument-based assurance cases asks reviewers to state the property at issue, specify the context, and connect the claim to evidence through an explicit warrant \parencite{icoAssuranceCases2026}. Recent AI safety-case work adapts that pattern to systems whose capabilities, deployment contexts, and risk profiles change over time \parencite{lee2026structuredSafetyCases}. Delegation assurance can be read as a reusable safety-case pattern for delegated machine action.

\textcite{rajagopalan2026governingAgenticAI} proposes an activity-based governance stack with six layers: registration, identification, verification, financial responsibility, lifecycle traceability, and suspension. These mechanisms supply an external enforcement framework by making responsibility-bearers reachable and enabling consequential activity to be stopped. They don't establish whether a concrete machine action was validly authorized or whether the recorded sequence from governing language to action can be reconstructed. Delegation assurance addresses that action-level question.

Governance and provenance standards supply adjacent evidence primitives. NIST's AI RMF and generative-AI profile foreground governance, monitoring, documentation, provenance, testing, and incident disclosure \parencite{nist2023artificial,nist2024generative}. W3C PROV, C2PA, and Certificate Transparency show how lineage, signed assertions, manifests, timestamps, hashes, and append-only logs can support reconstruction and tamper detection \parencite{moreau2013provdm,c2pa2026spec,laurie2021certificateTransparencyV2}.

Those mechanisms are necessary but normatively thin. Provenance can show where content came from, and an append-only log can make later alteration harder. They don't establish a string occurrence's contextual status, its source's standing, or whether \(R\) permits disclosure, durable-state update, or external action. Delegation assurance specifies the typed authorization claim and the fields needed to diagnose and compare performance. The evidentiary-assurance account \parencite{reynolds2026evidentiaryAssurance} asks whether the preserved record makes that claim evaluable and what evidentiary verdict it warrants.

\subsection{Evaluation and Automated-Evaluator Reliability}

Evaluation work supplies both tools and risks. LLM-as-judge methods can reduce grading cost and support scalable model comparison \parencite{zheng2023judging,liu2023geval}. Instruction-following benchmarks can make some compliance judgments more reproducible by using objectively checkable constraints \parencite{zhou2023ifeval}. These methods help build evidence about agent behaviour, but they don't automatically support an action-authorization claim.

Automated-evaluator reliability is itself part of the assurance problem. LLMBar shows that fluent but instruction-violating outputs can mislead evaluators \parencite{zeng2024llmbar}. Survey work treats consistency, bias, and domain adaptation as continuing problems for LLM-as-judge systems \parencite{gu2025judgeSurvey}. Bias studies show that both human and LLM judges can be vulnerable to judgment biases and attacks on those biases \parencite{chen2024judgementBias}. Steerability results show that, when supplied with a modified safety definition, LLM judges often continue to apply trained priors rather than the supplied definition \parencite{alloula2026safetyContextual}. If their labels feed downstream decisions, the operational rubric can diverge from the declared one.

Evaluation becomes part of the assurance chain when an institutional or system decision rule treats a label as evidence for a safety, compliance, or deployment claim. The record should identify who authorized that use, its scope, and the applicable decision rule. Evaluation methods contribute evidence only when the proposed interpretation and use of their labels has a target-matched validation argument.

\subsection{The Gap}

The gap is the coordinated assurance case those components jointly need to support: that the concrete operation was authorized and within scope, and that its trace supports reconstruction and attribution by an assurance reviewer.
