\section[Implications]{Implications for System Cards, Red-Teaming, Audits, and Automated-Evaluator Validation}

For system cards, delegation assurance changes the form of the claim. A task-suite pass rate or safe-final-answer rate leaves the projective claim underspecified. The card should declare the source result, outcome, evaluation bearer, unit, target cases, authority regime, allowed transformations, time and version range, tolerance, intended use, minimum useful reach, warrant, and prospective revision rule. It should report the normative references, recorded representations, item findings, and sample results separately and identify potentially answerable roles. Determining what the preserved record warrants requires the separate evidentiary-assurance analysis.

The useful artifact is a temporal delegation matrix. It records the action type and authority object; authority regime and higher-order basis; founding or adoption instrument and adoption competence; superior and non-delegable constraints; governing sources and effective intervals; standing to adopt, issue, direct, amend, suspend, revoke, override, review, or release; joint approvals and other guards; defect effects; permitted tools and data classes; persistence and disclosure permissions; expiry and emergency conditions; and each purported transition and its result.

Mean pass rates should be treated as descriptive summaries of the tested set and paired with item-level transitions and the prespecified worst-tail summaries defined in Section~\ref{sec:comparative-test}. Worst-tail degradation and absolute worst-tail loss answer different questions and shouldn't be collapsed. Oppositely signed context effects can cancel in an aggregate, and the affected items may be model-specific \parencite{zhang2026illusionRobustness}. A mean doesn't by itself support a deployment-wide assurance verdict. A system card should bind contextual-reliability results to the tested model and configuration and require revalidation after a material update.

Current-generation frontier system cards show the distinction. Anthropic describes system cards as documents for capabilities, safety evaluations, and responsible deployment decisions, and its Claude Opus 4.8 card reports extensive pre-deployment evaluations across responsible-scaling risks, cyber capabilities, safeguards and harmlessness, agentic safety, alignment, model welfare, and capabilities \parencite{anthropic2026modelSystemCards,anthropic2026opus48SystemCard}. That's the right level for model-level disclosure.

A deployment using such a model for delegated action would still need a delegation-specific record: which operation was authorized, which tool and data scopes applied, which people, organizations, or roles were competent to issue each relevant transition and which instruments supplied or recorded its authority, which information could be read but not disclosed, which persistent state changes were allowed, and which people and organizations occupy roles that may leave them answerable for the concrete action.

For red-teaming, delegation assurance separates attack success from evidentiary failure. One prompt-injection attempt may expose an item-level mismatch in recorded status, standing, or priority even when the final output is safe; a reliability claim requires repeated target-matched tests.

A tool-call proposal may be harmless in a test environment while exposing missing authorization at the gate. A refusal may reveal abandonment of required work, or only task failure when completion was permitted but not required. Red-team reports should preserve the attempted unauthorized redirection or proposed change to the recorded authorization state, the recorded classification or observable action-selection decision, the proposal, gate decision, delivery or execution outcome, and automated-evaluator label or expert adjudication rather than folding the event into a single pass/fail outcome.

For audits, reconstruction is a delegation-assurance evaluation endpoint. The authorization trace shouldn't rely on a model-generated statement that the action was authorized. It should expose governing sources, operation-specific standing, effective intervals, joint or guarded conditions, purported transitions and their recorded and adjudicated effects, provenance channel and content type, the applied priority rule, tool-call proposal, gate decision, executed operation, policy constraints, information-flow permissions, potentially answerable people and organizations, and state transitions. It should also name whether the regime is stipulated or institutional and preserve the prospective applicability declaration. That record becomes the input to evidentiary assurance; its completeness doesn't prove the asserted authority.

For a long-running trajectory, the same trace has to bind descendant sessions, retries, intermediate artefacts and state changes, monitor alerts, pause decisions, operational-continuation decisions, and the stopping condition. OpenAI reports adding trajectory monitoring and pause-and-review controls after a source conflict that culminated in sandbox circumvention and, separately, fragmented-token reconstruction; its account motivates these record fields but doesn't validate their effectiveness \parencite{openai2026longHorizonSafety}. A monitor alert is evidence and a control output, not an authorization verdict. The ability to pause execution isn't standing. Operational continuation changes the gate state, not \(Q_R\), unless it also constitutes a valid release event under \(R\) by a source with the relevant competence and with the applicable guards satisfied.

Execution through a model-based deployment doesn't make the model an independently reachable and answerable delegate. Operational attribution locates an event or failure among model action selection, scaffold or routing, policy-enforcement point, tool or runtime, and logging (Table~\ref{tab:delegation-evidence}). Institutional attribution identifies the people and organizations in roles that configured, deployed, approved, or monitored the system.

The trace also identifies who has to give reasons or respond and the answerability route through which that bearer can be contacted. That route needn't be the object-level review or remedy forum, whose access, competence, independence, and powers have to be assessed separately. These fields don't supply a remedy, confer unilateral standing to amend, revoke, or override a protected mandate, or determine legal liability. The governance record should document review and operational correction; financial recovery and lawful cessation depend on the external enforcement framework \parencite{rajagopalan2026governingAgenticAI}.

For architecture, delegation assurance warns against treating central coordination as free. Coase's theory of the firm makes the structural point: internal coordination can reduce transaction costs, but growth brings organizing costs and mistakes \parencite{coase1937nature}. The AI analogue is a coordinator agent with too many tools, too much context, and too much discretionary authority. Bounded subdelegations, local permissions, typed handoffs, and inspectable records can reduce the cost and complexity of assurance when they narrow or better delimit the permitted action set, localize governing constraints, and make the relevant record easier to reconstruct.

Some policy-enforcement points need a protected mandate so that a coordinator or deployer message can't unilaterally relax their constraints. Such protection supports an authorization claim only when it satisfies the higher-order validity, superior-constraint, review, and release conditions specified in Section~\ref{sec:delegation-beyond-responsiveness}.

For automated-evaluator assessment, evaluation belongs inside the assurance chain. A final-answer-only LLM-judge score doesn't establish whether an operation stayed authorized. Randomized local discrimination tests sensitivity to assigned representation packages. Predictive ablation tests incremental out-of-sample value. Reviewer reconstruction ablation tests what a configured reviewer population can warrant from typed versus degraded records. These programmes need separate results and validation arguments.

A delegation-assurance report should state the automated evaluator's assigned claim, who authorized its use, what evidence it received, how its labels gain operational effect, where expert adjudication remains necessary, and how uncertainty affects the assurance claim.
