\section{Conclusion}

Delegation assurance supplies regime-relative authorization semantics and typed representations for action and trajectory authorization, failure diagnosis, and comparative evaluation. These tools matter because AI deployments combine people, models, tools, persistent memory, and evaluators in ways that can obscure the distinctions among normative grants, recorded authorization states, proposed operations, executed trajectories, and assessment findings.

The framework treats the single action as the \(n=1\) trajectory case, represents local, prefix, cumulative, and terminal conditions, and distinguishes a normative hypergraph, recorded propagation graph, and execution trajectory. It also separates stipulated closed-world regimes from institutional regimes whose identity and applicability require adjudication. A record can support reconstruction without proving authority; an authorized action can be poorly recorded; and a pristine record can preserve an invalid grant. The executable artifact derives only the declared closed-world subset; it isn't an institutional-law engine.

Delegation assurance stakes its empirical usefulness on three separable tests: randomized local discrimination, predictive representation ablation, and reviewer reconstruction ablation. Their first specifications, input schemas, integrity rules, thresholds, and noncompensatory analyzers are frozen before target outcomes; only synthetic regression tests have been run. Every extension beyond observed cases requires a projective declaration, intended use, warrant plan, and prospective revision rule. The framework has no scalar measure of authority, and noncompensation applies only where the regime or validated assessment use is conjunctive. Its scope remains limited to cases where action permissions, limits, authority transitions, trajectories, or authorization assessment are at issue; it doesn't turn every model error into a delegation failure.
